---
title: "Data anaylsis, calibration and sensitivity"
author: "Rodrigo Souto-Veiga"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    title: "Data anaylsis, calibration and sensitivity"
execute: 
  warning: false
editor: 
  markdown: 
    wrap: 72
---

In this document are the new data analysis, calibration, and sensitivity
analysis for the model MetaSqueeze used in the following article:

**Souto-Veiga, R., Groeneveld, J., Enright, N. J., Fontaine, J. B., &
Jeltsch, F. (20XX). Climate change may shift metapopulations towards
unstable source-sink dynamics in a fire-killed, serotinous shrub**

Corresponding author:

Rodrigo Souto-Veiga^1, 2^

[rsoutoveiga\@uni-potsdam.de](rsoutoveiga@uni-potsdam.de)

ORCID: [0000-0001-8639-620X](https://orcid.org/0000-0001-8639-620X)

^1^ Plant Ecology and Nature Conservation, University of Potsdam, Am
Mühlenberg 3, 14476, Potsdam, Germany

^2^ Environmental and Conservation Sciences, Murdoch University, Murdoch
6150, WA, Australia


# 1. Initial cones and seeds

Calibration of initial cones per plant and viable seeds per cone. We
assumed that the plants are between 20 and 30 years old. Thus, we
calculate the average number of cones and average viable seeds per plant
between 20 and 30 years from 1000 replicates.

Mean number of cones per plant of 20 to 30 years after the fire mean =
85.868, SD = 4.309

Mean number of viable seeds per plant of 20 to 30 year after the fire
mean = 623.141, and SD = 22.060

```{r}
#| label: cali-1
#| include: true
#| echo: false
#| collapse: false
#| warning: false
#| eval: false

# load necessary packages
library(dplyr)

# relative path
g_file_path <- "data/sim_outputs/calibration/01_cali_init_cones_seeds/raw/"
g_file_name <- "01_cali_init_cones_seeds_sim_1_seed_dynamics.csv"

g_file_name <- base::paste0(g_file_path, g_file_name)

g_df <- utils::read.csv(g_file_name, sep = ',')

g_df_summ <- g_df %>% 
  dplyr::group_by(year) %>% 
  dplyr::summarise(
    mean_cones = base::mean(cones_plant),
    sd_cones   = stats::sd(cones_plant),
    mean_seeds = base::mean(seeds_plant),
    sd_seeds   = stats::sd(seeds_plant)
  )

# Mean number of cones per plant of 20 to 30 years after the fire
# mean = 85.868, SD = 4.309
base::print("number of cones per plant: mean and SD")
base::mean(g_df_summ$mean_cones[g_df_summ$year >= 20 & g_df_summ$year <= 30])
stats::sd(g_df_summ$mean_cones[g_df_summ$year >= 20 & g_df_summ$year <= 30])

# Mean number of viable seeds per plant of 20 to 30 year after the fire
# mean = 623.141, and SD = 22.060
base::print("number of viable seeds per plant: mean and SD")
base::mean(g_df_summ$mean_seeds[g_df_summ$year >= 20 & g_df_summ$year <= 30])
stats::sd(g_df_summ$mean_seeds[g_df_summ$year >= 20 & g_df_summ$year <= 30])

```

# 2. Prefire population size

## 2.1 Mortality scenarios

```{r}
#| label: mort-scn
#| include: true
#| echo: false
#| collapse: false
#| warning: false
#| eval: false

# load libraries
library(ggplot2)

g_file_path <- "data/papers/"
g_file_name <- "Keith2014_mort_rates.csv"

g_df <- read.csv(
  file = base::paste0(g_file_path, g_file_name), header = TRUE, sep = ','
)

# split spring and autumn data in two df
g_df_spring <- g_df[, c(1, 2, 3, 6)]
g_df_autumn <- g_df[, c(1, 2, 4, 6)]

# change moratlity rate column name
base::colnames(g_df_spring) <- c("year", "age", "mort", "lagrain")
base::colnames(g_df_autumn) <- c("year", "age", "mort", "lagrain")

# add column "season"
g_df_spring$season <- "spring"
g_df_autumn$season <- "autumn"

# fit morality rates from spring and autumn separately
g_lm_spring <- stats::lm(mort ~ I(1/age) + lagrain, data = g_df_spring)
g_lm_autumn <- stats::lm(mort ~ I(1/age) + lagrain, data = g_df_autumn)

# calculate predicted mortality rate based on plant age with mean lagrain
g_df_spring$y <- stats::coef(g_lm_spring)[1] +
  stats::coef(g_lm_spring)[2] * 1/g_df_spring$age +
  stats::coef(g_lm_spring)[3] * base::mean(g_df_spring$lagrain)

g_df_autumn$y <- stats::coef(g_lm_autumn)[1] +
  stats::coef(g_lm_autumn)[2] * 1/g_df_autumn$age +
  stats::coef(g_lm_autumn)[3] * base::mean(g_df_autumn$lagrain)

# combine df
g_df <- base::rbind(g_df_spring, g_df_autumn)

pmort <- ggplot(g_df) +
  geom_point(aes(x = age, y = mort, color = season), alpha = 0.5, size = 3) +
  geom_line(aes(x = age, y = y, color = season), size = 1.5) +
  theme_classic() +
  scale_color_manual(
    name = "Site",
    # values = c("#56B4E9", "#E69F00"),
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("autumn", "spring")
  ) +
  labs(x = "Plant age (years)", y = "Mortality rate") +
  ggplot2::scale_x_continuous(
    breaks = base::seq(from = 1, to = 15, by = 2)
  ) +
  theme(
    text = element_text(size = 14, face = "bold"),
    axis.text    = element_text(size = 14),
    axis.title   = element_text(size = 14, face = "bold"),
    legend.text  = element_text(size = 14, face = "plain"),
    legend.title = element_text(size = 14, face = "bold"),
    legend.position = c(0.8, 0.7)
  )

pmort

# save plot (uncomment code to save image)
ggplot2::ggsave(
  plot = pmort,
  filename = base::paste0("mort_fit_spring_autumn.png"),
  path = "./data/plots",
  width = 3.31 * 2,
  height = 4,
  bg = "white",
  dpi = 1200
)

```

```{r}
#| label: mort-plots
#| include: true
#| echo: false
#| collapse: false
#| warning: false
#| eval: false

# load libraries
library(ggplot2)
library(dplyr)
library(signs)
library(ggpubr)

# parameters used in the simulation experiments
g_plant_age_senescence <- 25
g_senescence_increase <- 0.01
g_long_term_rainfall <- 454
g_recruit_weather <- 0.06
g_recruit_mean <- 0.923
g_recruit_min <- 0.911
g_mort_min <- 0.02
g_mort_max <- 0.987

# Here I plot Enright et al. (1996) flower production and my new adjusted curve
g_df_mort_relative <- base::data.frame(age = base::seq(0, 40, 1))
g_df_mort_avg <- base::data.frame(age = base::seq(0, 40, 1))
g_df_mort_min <- base::data.frame(age = base::seq(0, 40, 1))

# functions
fun_mort_relative_age <- function(age) {
  base::ifelse(
    age == 0, 
    g_recruit_mean, 
    base::ifelse(
      age > 0 & age < 15, 
      0.0553 + (0.2645 / age) + (0.179 / age^2),
      base::ifelse(
        age >= 15 & age <= 25,
        0.0553 + (0.2645 / 15) + (0.179 / 15^2),
        (0.0553 + (0.2645 / 15) + (0.179 / 15^2)) + 
          (g_senescence_increase * (age - g_plant_age_senescence))
      )
    )
  )
}

fun_mort_spring <- function(age, rainfall) {
  (0.2877622 / age) + (-0.0005888 * rainfall) + 0.3179949
}

fun_mort_autumn <- function(age, rainfall) {
  (0.6475991 / age) + (-0.0006502 * rainfall) + 0.2867799
}

fun_mort_relative_weather <- function(age, rainfall) {
  base::ifelse(
    age == 0,
    1 + (g_long_term_rainfall - rainfall) / 100 * g_recruit_weather, 
    1 + (-0.00061 * rainfall + 0.26873))
}

fun_mort_average <- function(age, rainfall) {
  base::ifelse(
    age == 0, 
    g_recruit_mean + ((g_long_term_rainfall - rainfall) / 100 * g_recruit_weather), 
    base::ifelse(
      age > 0 & age <= 15, 
      (fun_mort_spring(age, rainfall) + fun_mort_autumn(age, rainfall)) / 2,
      base::ifelse(
        age > 15 & age <= 25,
        (fun_mort_spring(15, rainfall) + fun_mort_autumn(15, rainfall)) / 2,
        (fun_mort_spring(15, rainfall) + fun_mort_autumn(15, rainfall)) / 2 + 
        (g_senescence_increase * (age - g_plant_age_senescence))
      )
    )
  )
}

fun_mort_min <- function(age, rainfall) {
  base::ifelse(
    age == 0, 
    g_recruit_min + ((g_long_term_rainfall - rainfall) / 100 * g_recruit_weather), 
    base::ifelse(
      age > 0 & age <= 15, 
      base::min(fun_mort_spring(age, rainfall), fun_mort_autumn(age, rainfall)),
      base::ifelse(
        age > 15 & age <= 25,
        base::min(fun_mort_spring(15, rainfall), fun_mort_autumn(15, rainfall)),
        base::min(fun_mort_spring(15, rainfall), fun_mort_autumn(15, rainfall)) + 
          (g_senescence_increase * (age - g_plant_age_senescence))
      )
    )
  )
}

fun_truncate_mortality <- function(age, mort) {
  base::ifelse(
    mort > g_mort_max, 
    g_mort_max, 
    base::ifelse(
      age == 0 & mort < g_recruit_min,
      g_recruit_min,
      base::ifelse(
        mort < g_mort_min,
        g_mort_min,
        mort
      )
    )
  )
}

# scenario name
g_df_mort_relative$scn <- "relative"
g_df_mort_avg$scn <- "average"
g_df_mort_min$scn <- "min"

# calculate moratlity rates values for mort relative function
g_df_mort_relative_long <- g_df_mort_relative

g_df_mort_relative_long <- g_df_mort_relative_long %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_relative_age(age))

g_df_mort_relative_long <- g_df_mort_relative_long %>% 
  group_by(age, scn) %>%
  mutate(mort = mort * fun_mort_relative_weather(age, 454))

g_df_mort_relative_long$rain <- "454"

# +100 mm
g_df_mort_relative_plus100 <- g_df_mort_relative
g_df_mort_relative_plus100 <- g_df_mort_relative_plus100 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_relative_age(age))

g_df_mort_relative_plus100 <- g_df_mort_relative_plus100 %>% 
  group_by(age, scn) %>%
  mutate(mort = mort * fun_mort_relative_weather(age, 554))

g_df_mort_relative_plus100$rain <- "554"

# +200 mm
g_df_mort_relative_plus200 <- g_df_mort_relative
g_df_mort_relative_plus200 <- g_df_mort_relative_plus200 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_relative_age(age))

g_df_mort_relative_plus200 <- g_df_mort_relative_plus200 %>% 
  group_by(age, scn) %>%
  mutate(mort = mort * fun_mort_relative_weather(age, 654))

g_df_mort_relative_plus200$rain <- "654"

# -100 mm
g_df_mort_relative_minus100 <- g_df_mort_relative
g_df_mort_relative_minus100 <- g_df_mort_relative_minus100 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_relative_age(age))

g_df_mort_relative_minus100 <- g_df_mort_relative_minus100 %>% 
  group_by(age, scn) %>%
  mutate(mort = mort * fun_mort_relative_weather(age, 354))

g_df_mort_relative_minus100$rain <- "354"

# -200 mm
g_df_mort_relative_minus200 <- g_df_mort_relative
g_df_mort_relative_minus200 <- g_df_mort_relative_minus200 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_relative_age(age))

g_df_mort_relative_minus200 <- g_df_mort_relative_minus200 %>% 
  group_by(age, scn) %>%
  mutate(mort = mort * fun_mort_relative_weather(age, 254))

g_df_mort_relative_minus200$rain <- "254"


g_df_mort_relative_all <- base::rbind(
  g_df_mort_relative_plus200,
  g_df_mort_relative_plus100,
  g_df_mort_relative_long,
  g_df_mort_relative_minus100,
  g_df_mort_relative_minus200
)

# truncate mortality probability
g_df_mort_relative_all <- g_df_mort_relative_all %>% 
  mutate(
    mort = fun_truncate_mortality(age, mort)
  )

g_df_mort_relative_all$rain

g_df_mort_relative_all$rain <- base::as.factor(g_df_mort_relative_all$rain)
g_df_mort_relative_all$rain <- base::ordered(g_df_mort_relative_all$rain)

p1 <- ggplot(data = g_df_mort_relative_all,
       aes(x = age,
           y = mort,
           color = rain)) +
  geom_line(size = 1.2,
            alpha = 0.7) +
  scale_color_manual(
    name = "Rainfall scenario",
    values = c("#D55E00", "#E69F00", "black", "#56B4E9", "#0072B2"),
    labels = c(
      "\u2013200 mm below average",
      "\u2013100 mm below average",
      "long-term average [454 mm]",
      "+100 mm above average",
      "+200 mm above average"
    )
  ) +
  theme_classic() +
  xlab("Plant age (years)") +
  ylab("Mortality") +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_mort_max,
      linetype = "max plant mortality"),
    color = "black"
  ) +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_recruit_min,
      linetype = "min recruit mortality"),
    color = "black"
  ) +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_mort_min,
      linetype = "min plant mortality"),
    color = "black"
  ) +
  theme(
    text = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  ) +
  ggplot2::scale_linetype_manual(
    name = "",
    breaks = c(
      "max plant mortality", 
      "min recruit mortality", 
      "min plant mortality"),
    values = c(2, 3, 4)
  ) +
  ggplot2::ggtitle("(a)")


#############################################
g_df_mort_avg_long <- g_df_mort_avg

g_df_mort_avg_long <- g_df_mort_avg_long %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_average(age, 454))

g_df_mort_avg_long$rain <- "454"

# +100 mm
g_df_mort_avg_plus100 <- g_df_mort_avg
g_df_mort_avg_plus100 <- g_df_mort_avg_plus100 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_average(age, 554))

g_df_mort_avg_plus100$rain <- "554"

# +200 mm
g_df_mort_avg_plus200 <- g_df_mort_avg
g_df_mort_avg_plus200 <- g_df_mort_avg_plus200 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_average(age, 654))

g_df_mort_avg_plus200$rain <- "654"

# -100 mm
g_df_mort_avg_minus100 <- g_df_mort_avg
g_df_mort_avg_minus100 <- g_df_mort_avg_minus100 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_average(age, 354))

g_df_mort_avg_minus100$rain <- "354"

# -200 mm
g_df_mort_avg_minus200 <- g_df_mort_avg
g_df_mort_avg_minus200 <- g_df_mort_avg_minus200 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_average(age, 254))

g_df_mort_avg_minus200$rain <- "254"

g_df_mort_avg_all <- base::rbind(
  g_df_mort_avg_plus200,
  g_df_mort_avg_plus100,
  g_df_mort_avg_long,
  g_df_mort_avg_minus100,
  g_df_mort_avg_minus200
)

# truncate mortality probability
g_df_mort_avg_all <- g_df_mort_avg_all %>% 
  mutate(
    mort = fun_truncate_mortality(age, mort)
  )



g_df_mort_avg_all$rain <- base::as.factor(g_df_mort_avg_all$rain)
g_df_mort_avg_all$rain <- base::ordered(g_df_mort_avg_all$rain)


p2 <- ggplot(data = g_df_mort_avg_all,
       aes(x = age,
           y = mort,
           color = rain)) +
  geom_line(size = 1.2,
            alpha = 0.7) +
  scale_color_manual(
    name = "Rainfall scenario",
    values = c("#D55E00", "#E69F00", "black", "#56B4E9", "#0072B2"),
    labels = c(
      "\u2013200 mm below average",
      "\u2013100 mm below average",
      "long-term average [454 mm]",
      "+100 mm above average",
      "+200 mm above average"
    )
  ) +
  theme_classic() +
  xlab("Plant age (years)") +
  ylab("Mortality") +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_mort_max,
      linetype = "max plant mortality"),
    color = "black"
  ) +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_recruit_min,
      linetype = "min recruit mortality"),
    color = "black"
  ) +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_mort_min,
      linetype = "min plant mortality"),
    color = "black"
  ) +
  theme(
    text = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  ) +
  ggplot2::scale_linetype_manual(
    name = "",
    breaks = c(
      "max plant mortality", 
      "min recruit mortality", 
      "min plant mortality"),
    values = c(2, 3, 4)
  ) +
  ggplot2::ggtitle("(b)")


###############################################
g_df_mort_min_long <- g_df_mort_min

g_df_mort_min_long <- g_df_mort_min_long %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_min(age, 454))

g_df_mort_min_long$rain <- "454"

# +100 mm
g_df_mort_min_plus100 <- g_df_mort_min
g_df_mort_min_plus100 <- g_df_mort_min_plus100 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_min(age, 554))

g_df_mort_min_plus100$rain <- "554"

# +200 mm
g_df_mort_min_plus200 <- g_df_mort_min
g_df_mort_min_plus200 <- g_df_mort_min_plus200 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_min(age, 654))

g_df_mort_min_plus200$rain <- "654"

# -100 mm
g_df_mort_min_minus100 <- g_df_mort_min
g_df_mort_min_minus100 <- g_df_mort_min_minus100 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_min(age, 354))

g_df_mort_min_minus100$rain <- "354"

# -200 mm
g_df_mort_min_minus200 <- g_df_mort_min
g_df_mort_min_minus200 <- g_df_mort_min_minus200 %>% 
  group_by(age, scn) %>%
  mutate(mort = fun_mort_min(age, 254))

g_df_mort_min_minus200$rain <- "254"

g_df_mort_min_all <- base::rbind(
  g_df_mort_min_plus200,
  g_df_mort_min_plus100,
  g_df_mort_min_long,
  g_df_mort_min_minus100,
  g_df_mort_min_minus200
)

# truncate mortality probability
g_df_mort_min_all <- g_df_mort_min_all %>% 
  mutate(
    mort = fun_truncate_mortality(age, mort)
  )



g_df_mort_min_all$rain <- base::as.factor(g_df_mort_min_all$rain)
g_df_mort_min_all$rain <- base::ordered(g_df_mort_min_all$rain)

p3 <- ggplot(data = g_df_mort_min_all,
       aes(x = age,
           y = mort,
           color = rain)) +
  geom_line(size = 1.2,
            alpha = 0.7) +
  scale_color_manual(
    name = "Rainfall scenario",
    values = c("#D55E00", "#E69F00", "black", "#56B4E9", "#0072B2"),
    labels = c(
      "\u2013200 mm below average",
      "\u2013100 mm below average",
      "long-term average [454 mm]",
      "+100 mm above average",
      "+200 mm above average"
    )
  ) +
  theme_classic() +
  xlab("Plant age (years)") +
  ylab("Mortality") +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_mort_max,
      linetype = "max plant mortality"),
    color = "black"
  ) +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_recruit_min,
      linetype = "min recruit mortality"),
    color = "black"
  ) +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_mort_min,
      linetype = "min plant mortality"),
    color = "black"
  ) +
  theme(
    text = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  ) +
  ggplot2::scale_linetype_manual(
    name = "",
    breaks = c(
      "max plant mortality", 
      "min recruit mortality", 
      "min plant mortality"),
    values = c(2, 3, 4)
  ) +
  ggplot2::ggtitle("(c)")



pleyend <- ggplot() +
  geom_line(
    data = g_df_mort_min_all,
    mapping = ggplot2::aes(
      x = age,
      y = mort,
      color = rain), 
    size = 1.2,
    alpha = 0.7) +
  scale_color_manual(
    name = "Rainfall scenario",
    values = c("#D55E00", "#E69F00", "black", "#56B4E9", "#0072B2"),
    labels = c(
      "\u2013200 mm below average",
      "\u2013100 mm below average",
      "long-term average [454 mm]",
      "+100 mm above average",
      "+200 mm above average"
    )
  ) +
  ggplot2::xlim(0, 0) +
  ggplot2::ylim(0, 0) +
  ggtitle(" ") +
  theme(
    legend.position = c(0.5, 0.5),
    legend.text  = element_text(size = 28, face = "plain"),
    legend.title = element_text(size = 28, face = "bold"),
    legend.key.size = unit(3, "line")
  ) +
  # guides(colour = guide_legend(override.aes = list(size = 9))) +
  theme_void() +
  xlab("Plant age (years)") +
  ylab("Annual mortality probability") +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_mort_max,
      linetype = "max plant mortality"),
    color = "black"
  ) +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_recruit_min,
      linetype = "min recruit mortality"),
    color = "black"
  ) +
  geom_hline(
    mapping = ggplot2::aes(
      yintercept = g_mort_min,
      linetype = "min plant mortality"),
    color = "black"
  ) +
  ggplot2::scale_linetype_manual(
    name = "",
    breaks = c(
      "max plant mortality", 
      "min recruit mortality", 
      "min plant mortality"),
    values = c(2, 3, 4)
  )


g_parr <- ggpubr::ggarrange(
  p1, p2, p3, pleyend, 
  ncol = 2,
  nrow = 2
)

g_parr

# save plot (uncomment code to save image)
ggplot2::ggsave(
  plot = g_parr,
  filename = base::paste0("mort_scenario_curves_NEW_dpi300.tiff"),
  path = "./data/plots",
  width = 3.31 * 2,
  height = 3.31 * 2,
  bg = "white",
  dpi = 300
)

```

The initial number of plants in each occupied dune.

located in data/calibration/02_cali

```{r}
#| label: dune
#| include: true
#| echo: false
#| collapse: false
#| warning: false
#| eval: false

# load necessary packages
library(dplyr)

# read dune data from previous papers (He et al., 2004, 2010)
g_file_path <- "data/papers/"
g_file_name <- "He2004_2010_dune_data.csv"

g_file_name <- base::paste0(g_file_path, g_file_name)

g_dunes <- utils::read.csv(g_file_name, sep = ',')

g_dunes <- g_dunes %>% 
  dplyr::filter(!base::is.na(code_he2010))

# first estimate of prefire pop size in NA dunes
g_dunes <- g_dunes %>% 
  dplyr::mutate(
    prefire_pop = base::ifelse(
      base::is.na(prefire_pop_size_he2004),
      100,
      prefire_pop_size_he2004
    ) 
  )

g_dunes <- g_dunes %>% 
  dplyr::mutate(pop_size_he2010 = dune_area_ha_he2010 * pop_density_he2010)

g_dunes <- g_dunes[base::order(g_dunes$code_he2010), ]
```

Calibration of prefire population for each mortality scenario keeping
linear relationship between prefire population size and predicted mean
population size in 2006 (8 year-old plants). I used the so-called
"cross-multiplication" method
https://en.wikipedia.org/wiki/Cross-multiplication

So we can say that the new prefire population size was calculated using
the cross-multiplication knowing the first estimated prefire population,
the observed population size in 2006 and the predicted population size
in 2006.

NOTE: the mortality scenarios 3 and 4 have the same number of initial
plants (i.e. prefire population size), because there is no LDD (and thus
no immigrants)

```{r}
#| label: cali-2-prefire_pop_size
#| include: true
#| echo: false
#| collapse: false
#| warning: false
#| eval: false
#| cache: true
#| cache-globals: g_dunes

# load libraries
library(plyr)
library(dplyr)

# load utils
base::source("utils/read_multiple_files.R")

# read calibration 02
# initial prefire population size for each mortality scenario
# mort scn 0: age-weather mortality curves (Souto-Veiga et al., 2022)
# mort scn 2: spring and autumn mortality curves, arithmetic mean
# mort scn 3: LDD cohorts (spring) and SDD cohorts (autumn)

g_file_path <- "data/sim_outputs/calibration/02_01_cali_prefire_pop/raw"
g_sim_type <- "*seed_dynamics.csv"

# read files and create df
g_df <- fun_read_files(g_file_path, g_sim_type)

# population size in 2006 (8-year-old plants) for each mortality scenario
g_df <- g_df %>% 
  dplyr::filter(year == 8) %>% 
  dplyr::group_by(sim_id, pop) %>% 
  dplyr::summarise(
    mean_plants = base::round(base::mean(num_plants), 2),
    sd_cones   = base::round(stats::sd(num_plants), 2)
  )

# merge calculated data with observed data
g_df <- base::merge(
  x = g_dunes, 
  y = g_df, 
  by.x = "code_he2010", 
  by.y = "pop"
)

# take only dunes from He et al., 2010
g_df <- g_df %>% 
  dplyr::filter(!base::is.na(code_he2010))

# difference between calculated vs observed population size
g_df <- g_df %>% 
  dplyr::mutate(error_abs = mean_plants - pop_size_he2010)

g_df <- g_df %>% 
  dplyr::mutate(
    error_pct = base::round(
      base::abs(mean_plants - pop_size_he2010) / pop_size_he2010 * 100,
      2
    )
  )

# calculate new prefire population size for each mortality scenario
g_df <- g_df %>% 
  dplyr::mutate(
    new_prefire_pop = base::round(prefire_pop * pop_size_he2010 / mean_plants)
)

# order df by mortality scn and population id
g_df <- g_df[base::order(g_df$sim_id, g_df$code_he2010), ]

g_df_small <- g_df[, c(10, 1, 9, 11, 12, 13, 14, 6, 15)]

# change colnames of all columns
base::colnames(g_df_small) <- c(
  "sim id", 
  "pop id", 
  "observed plants", 
  "mean plants",
  "sd plants",
  "abs error",
  "pct error",
  "prefire pop",
  "new prefire pop"
)

knitr::kable(
  g_df_small, 
  caption = "Table 1. Calibration of prefire population for each mortality 
  scenario"
)


# utils::write.csv(x = g_df_small, file = "foo.csv")
```

Check the observed vs predicted population sizes in 2006 (8 year-old
plants)

The value above 10% error is in the row number 23, with 17.85% error.
This is because the population sizes are too small, i.e. 13 plants and
10.68 plants (observed and predicted mean).

```{r}
#| label: cali-2-prefire-pop-size-check
#| include: true
#| echo: false
#| collapse: false
#| warning: false
#| eval: false
#| cache: true
#| cache-globals: g_dunes

# load libraries
library(plyr)
library(dplyr)

# load utils
base::source("utils/read_multiple_files.R")

# read calibration 02 check
# initial prefire population size for each mortality scenario
# mort scn 0: age-weather mortality curves (Souto-Veiga et al., 2022)
# mort scn 2: spring and autumn mortality curves, arithmetic mean
# mort scn 3: LDD cohorts (spring) and SDD cohorts (autumn)

g_file_path <- "data/sim_outputs/calibration/03_cali_init_plants_mort_scn_check/raw"
g_sim_type <- "*seed_dynamics.csv"

# read files and create df
g_df <- fun_read_files(g_file_path, g_sim_type)

# population size in 2006 (8-year-old plants) for each mortality scenario
g_df <- g_df %>% 
  dplyr::filter(year == 8) %>% 
  dplyr::group_by(sim_id, pop) %>% 
  dplyr::summarise(
    mean_plants = base::round(base::mean(num_plants), 2),
    sd_cones = base::round(stats::sd(num_plants), 2)
  )

# merge calculated data with observed data
g_df <- base::merge(
  x = g_dunes, 
  y = g_df, 
  by.x = "code_he2010", 
  by.y = "pop"
)

# take only dunes from He et al., 2010
g_df <- g_df %>% 
  dplyr::filter(!base::is.na(code_he2010))

# difference between calculated vs observed population size
g_df <- g_df %>% 
  dplyr::mutate(error_abs = mean_plants - pop_size_he2010)

g_df <- g_df %>% 
  dplyr::mutate(
    error_pct = base::round(
      base::abs(mean_plants - pop_size_he2010) / pop_size_he2010 * 100,
      2
    )
  )

# order df by mortality scn and population id
g_df <- g_df[base::order(g_df$sim_id, g_df$code_he2010), ]

g_df_small <- g_df[, c(10, 1, 9, 11, 12, 13, 14)]

# change colnames of all columns
base::colnames(g_df_small) <- c(
  "sim id", 
  "pop id", 
  "observed plants", 
  "mean plants",
  "sd plants",
  "abs error",
  "pct error"
)

knitr::kable(
  g_df_small,
  caption = "Table 2. Check the percentage error of the calibration of prefire 
  pop size calculated in Table 1"
)
```

# 3. LDD

## 3.1 Dispersal kernel of seeds by postfire wind

```{r}
#| label: dispersal-kernel-wind
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)
library(fitdistrplus)
library(knitr)

# read dune data from previous papers (He et al., 2004, 2010)
g_file_path <- "data/papers/"
g_file_name <- "He2010_table3_immigrants.csv"

g_file_name <- base::paste0(g_file_path, g_file_name)

g_df <- utils::read.csv(g_file_name, sep = ',')

# get columns distance_km, remove NAs and convert to meters
g_df <- base::as.data.frame(g_df$distance_km)
g_df <- stats::na.omit(g_df)
base::colnames(g_df)[1] <- "distance_m"
g_df$distance_m <- g_df$distance_m * 1000

# histogram (density scale) and CDF. The distribution is unimodal (one peak)
# and skewed to the right, as expected
fitdistrplus::plotdist(g_df$distance_m, histo = TRUE, demp = TRUE)

# skewness-kurtosis plot. The distribution is highly skewed to the right 
# estimated skewness 1.31. The kurtoses is > 3 (5.17), thus leptokurtic
fitdistrplus::descdist(g_df$distance_m, boot = 1000)

# fitting pdf. Weibull, gamma, and log-norm
g_fw <- fitdistrplus::fitdist(g_df$distance_m, "weibull")
g_fg <- fitdistrplus::fitdist(g_df$distance_m, "gamma")
g_fln <- fitdistrplus::fitdist(g_df$distance_m, "lnorm")

# model selection based on AIC values. The three distributions show very
# similar values. The best (lowers AIC value) is the lognorm distritbution.
# Furthermore, lognorm is typically used for LDD, since outperforms many other
# distributions (Nathan et al., 2012). Also, Frank Schurr recommended lognorm
# summary(g_fw)
# summary(g_fln)
# summary(g_fg)

# get aic values for each distribution
g_fw_aic <- base::data.frame(distribution = "weibull", AIC = g_fw$aic)
g_fg_aic <- base::data.frame(distribution = "gamma", AIC = g_fg$aic)
g_fln_aic <- base::data.frame(distribution = "lognorm", AIC = g_fln$aic)

g_aic <- base::rbind(g_fw_aic, g_fg_aic, g_fln_aic)

knitr::kable(g_aic, caption = "AIC values of the three distribution tested")


graphics::par(mfrow = c(2, 2))
plot.legend <- c("weibull", "gamma", "lognormal")
fitdistrplus::denscomp(list(g_fw, g_fg, g_fln), legendtext = plot.legend)
fitdistrplus::qqcomp(list(g_fw, g_fg, g_fln), legendtext = plot.legend)
fitdistrplus::cdfcomp(list(g_fw, g_fg, g_fln), legendtext = plot.legend)
fitdistrplus::ppcomp(list(g_fw, g_fg, g_fln), legendtext = plot.legend)

base::print("Selected dispersal kernel:")

# SVG graphics device
svg("./data/plots/wind_density_kernel.svg")

# plot
graphics::par(mfrow = c(1, 1), ps = 18)
graphics::hist(
  g_df$distance_m, breaks = 11, prob = TRUE,
  xlab = "Distance (m)",
  ylab = "Density",
  size = 3,
  xlim = c(0, 3500),
  ylim = c(0, 0.0015),
  xaxis = "i",
  yaxis = "i",
  main = ""
)
graphics::lines(
  dlnorm(base::seq(0, 3500, 1), meanlog = 6.7872, sdlog = 0.6756), 
  lwd = 3, col = "#0073C2FF"
)

# close the graphics device
dev.off()

```

## 3.2 Number of seeds per dispersed cone

Fit the number of follicles per cone to estimate the distribution

```{r}
#| label: follicles-cone-baseline
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)
library(fitdistrplus)
library(knitr)

# read dune data from previous papers (He et al., 2004, 2010)
g_file_path <- "data/papers/"
g_file_name <- "eneabba_follicles_per_cone.csv"

g_file_name <- base::paste0(g_file_path, g_file_name)

g_df <- utils::read.csv(g_file_name, sep = ',')

# histogram (density scale) and CDF.
fitdistrplus::plotdist(g_df$X1986, histo = TRUE, demp = TRUE)

# skewness-kurtosis plot.
fitdistrplus::descdist(g_df$X1986, boot = 1000)

# fitting most common discrete denstiy functions: Poisson, nbinom, and geom
fit_p <- fitdistrplus::fitdist(g_df$X1986, "pois")
fit_nb <- fitdistrplus::fitdist(g_df$X1986, "nbinom")
fit_g <- fitdistrplus::fitdist(g_df$X1986, "geom")

# get aic values for each distribution
fit_p_aic <- data.frame(distribution = "poisson", AIC = fit_p$aic)
fit_nb_aic <- data.frame(distribution = "negative binomial", AIC = fit_nb$aic)
fit_g_aic <- data.frame(distribution = "geometric", AIC = fit_g$aic)

g_aic <- base::rbind(
  fit_p_aic, 
  fit_nb_aic, 
  fit_g_aic
)

knitr::kable(g_aic, caption = "AIC values of the three distribution tested")

graphics::par(mfrow = c(2, 2))
plot.legend <- c("poisson", "negative binomial", "geometric")
fitdistrplus::denscomp(list(fit_p, fit_nb, fit_g), legendtext = plot.legend)
fitdistrplus::qqcomp(list(fit_p, fit_nb, fit_g), legendtext = plot.legend)
fitdistrplus::cdfcomp(list(fit_p, fit_nb, fit_g), legendtext = plot.legend)
fitdistrplus::ppcomp(list(fit_p, fit_nb, fit_g), legendtext = plot.legend)

# base::print("Selected dispersal kernel:")
# 
# graphics::par(mfrow = c(1, 1), ps = 18)
# graphics::hist(
#   g_df$X1986, breaks = 11, prob = TRUE,
#   xlab = "# of follicles per cone",
#   ylab = "density",
#   main = "Lognorm: meanlog = 6.787, sdlog = 0.676",
#   size = 3
# )
# graphics::barplot(
#   dnbinom(base::seq(0, 20, 1), size = fit_nb$estimate[1], mu = fit_nb$estimate[2]), 
#   lwd = 2, col = "#0073C2FF"
# )

g_df_follicles_base <- base::as.data.frame(g_df$X1986)
g_df_follicles_current <- base::as.data.frame(g_df$X2018)

g_df_follicles_current <- stats::na.omit(g_df_follicles_current)



g_df_follicles_base$scenario <- "Baseline"
g_df_follicles_current$scenario <- "Current"

base::colnames(g_df_follicles_base) <- c("follicles", "scenario")
base::colnames(g_df_follicles_current) <- c("follicles", "scenario")


g_df_follicles_all <- base::rbind(g_df_follicles_base, g_df_follicles_current)

p_follicles_all <- ggplot(
  data = g_df_follicles_all,
  mapping = aes(
    x = scenario,
    y = follicles,
    fill = scenario)) +
  geom_boxplot(
    alpha = 0.6,
    outlier.shape = 1,
    notch = TRUE,
    notchwidth = 0.5
  ) +
  geom_jitter(
    width = 0.2,
    size = 1,
    alpha = 0.4,
    shape = 21,
    color = "black",
    fill = "black"
  ) + 
  scale_fill_manual(    
    name = "Scenario",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("Baseline", "Current")
  ) +
  guides(fill = FALSE) +
  theme_classic() +
  xlab("Climate scenarios") +
  ylab("# of follicles per cone") +
  scale_y_continuous(
    breaks = base::seq(from = 0, to = 25, by = 5)
  ) +
    ggplot2::coord_cartesian(ylim = c(0, 28)) +
  theme(text = element_text(size = 12, face = "bold"),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12, face = "bold"),
        legend.text  = element_text(size = 12),
        legend.title = element_text(size = 12, face = "bold")) +
  ggsignif::geom_signif(test = "wilcox.test", comparisons = list(c(1, 2)), 
                        map_signif_level = TRUE, y_position = 25) +
  ggtitle("(a)")
# labs(subtitle = rstatix::get_test_label(stat_test, detailed = TRUE))


```

```{r}
#| label: follicles-cone-current
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)
library(fitdistrplus)
library(knitr)

# read dune data from previous papers (He et al., 2004, 2010)
g_file_path <- "data/papers/"
g_file_name <- "eneabba_follicles_per_cone.csv"

g_file_name <- base::paste0(g_file_path, g_file_name)

g_df <- utils::read.csv(g_file_name, sep = ',')

g_df <- na.omit(g_df)

# histogram (density scale) and CDF
fitdistrplus::plotdist(g_df$X2018, histo = TRUE, demp = TRUE)

# skewness-kurtosis plot
fitdistrplus::descdist(g_df$X2018, boot = 1000)

# fitting most common discrete denstiy functions: Poisson, nbinom, and geom
fit_p <- fitdistrplus::fitdist(g_df$X2018, "pois")
fit_nb <- fitdistrplus::fitdist(g_df$X2018, "nbinom")
fit_g <- fitdistrplus::fitdist(g_df$X2018, "geom")

# get aic values for each distribution
fit_p_aic <- data.frame(distribution = "poisson", AIC = fit_p$aic)
fit_nb_aic <- data.frame(distribution = "negative binomial", AIC = fit_nb$aic)
fit_g_aic <- data.frame(distribution = "geometric", AIC = fit_g$aic)

g_aic <- base::rbind(
  fit_p_aic, 
  fit_nb_aic, 
  fit_g_aic
)

knitr::kable(g_aic, caption = "AIC values of the three distribution tested")


graphics::par(mfrow = c(2, 2))
plot.legend <- c("poisson", "negative binomial", "geometric")
fitdistrplus::denscomp(list(fit_p, fit_nb, fit_g), legendtext = plot.legend)
fitdistrplus::qqcomp(list(fit_p, fit_nb, fit_g), legendtext = plot.legend)
fitdistrplus::cdfcomp(list(fit_p, fit_nb, fit_g), legendtext = plot.legend)
fitdistrplus::ppcomp(list(fit_p, fit_nb, fit_g), legendtext = plot.legend)

# base::print("Selected dispersal kernel:")
# 
# graphics::par(mfrow = c(1, 1), ps = 18)
# graphics::barplot(
#   dpois(base::seq(0, 20, 1), lambda = fit_p$estimate[1]), 
#   lwd = 2, col = "#EFC000FF"
# )

```

```{r}
#| label: follicles-in-dispersed-cones
#| include: true
#| echo: false
#| collapse: false
#| warning: false

library(tidyr)
library(ggplot2)

# the first two hours the canopy is divided by half due to the heat of the fire
# this is taken also the old cones, however, birds will go for the
# younger cones because they have more seeds and are more accessible

g_n = 0:26

g_df <- data.frame(
  events = g_n, 
  baseline = dnbinom(x = g_n, size = 6.219, mu = 10.08),
  current = dpois(x = g_n, lambda = 7.325)
)

# pivot data from wide to long
g_df_tidy <- g_df %>% 
  tidyr::pivot_longer(
    -c(events),
    names_to = "Scenario",
    values_to = "Density"
  )

# plot
p_follicles_distr <- ggplot(
  data = g_df_tidy, 
  mapping = aes(
    x = factor(events),
    y = Density,
    fill = Scenario)) +
  geom_col(position = "dodge") +
  labs(x = "# of follicles per cone", y = "Density") +
  theme_classic() +
  scale_fill_manual(
    name = "Scenario:",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("Baseline", "Current")
  ) +
  scale_x_discrete(breaks = seq(0, 26, 4)) +
  ggplot2::theme(
    text = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    strip.text.x = element_blank(),
    strip.background = element_rect(colour = "white", fill = "white"),
    legend.position = c(0.85, 0.7)
  ) +
  ggtitle("(b)")



g_parr <- ggpubr::ggarrange(
  p_follicles_all,
  p_follicles_distr,
  ncol = 1,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom"
)



# save plot (uncomment code to save image)
ggplot2::ggsave(
  plot = g_parr,
  filename = base::paste0("follicles_distr_a_b_dpi300.tiff"),
  path = "./data/plots",
  width = 3.31,
  height = 3.31 * 2,
  bg = "white",
  dpi = 300
)

```

## 3.3 Calibration of proportion of dispersed cones and seeds

Calibratrion criteria is 10% error of the observed immigration
percentage and average number of populations per dune in He et al.,
(2010)

```{r}
#| label: cali-3
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(plyr)
library(dplyr)
# library(easycsv)
library(ggplot2)
library(cowplot)
library(hrbrthemes)
library(viridis)

# load utils
base::source("utils/pooled_sd.R")
base::source("utils/read_multiple_files.R")

# input files
g_path <- "data/sim_outputs/calibration/03_cali"
g_sim_type <- "*metapop_dynamics.csv"

# output figure
g_figure_name <- "03_cali_perc_10_years_8_NEW_dpi1200"

# calibration reference values
g_percentage <- 10
g_immi_min <- 5.5
g_immi_max <- 6.8
g_pop_max <- 4

# read files and create df
g_df <- fun_read_files(paste0(g_path, "/raw"), g_sim_type)

##############################################
## read from zip. It takes too long
# easycsv::fread_zip(
#   base::paste0(`g_path, g_file, ".zip"), 
#   extension = "CSV", 
#   sep = ',', 
#   data.table = FALSE
# )
# 
# g_df <- do.call(rbind, mget(ls(pattern = g_file)))
# 
# # remove them from global environment
# rm(list = ls(pattern = g_file))
#############################################

# population size in 2006 (8-year-old plants) for each mortality scenario
g_df_immi_perc <- g_df %>% 
  dplyr::filter(year == 8) %>%
  dplyr::group_by(sim_id, study_num) %>% 
  dplyr::summarise(
    mean_immi_perc = base::mean(immi_perc),
    sd_immi_perc = stats::sd(immi_perc)
)

g_df_immi <- g_df %>% 
  # dplyr::filter(year == 5 | year == 8) %>%
  dplyr::filter(year == 8) %>%
  dplyr::group_by(sim_id, study_num) %>% 
  dplyr::summarise(
    mean_immi = base::mean(immi),
    sd_immi = stats::sd(immi)
)

g_df_immi_perc <- g_df_immi_perc %>% 
  dplyr::group_by(sim_id) %>% 
  dplyr::summarise(
    grand_mean_immi_perc = base::mean(mean_immi_perc),
    grand_sd_immi_perc = fun_pooled_sd_equal_sample_sizes(sd_immi_perc)
  )

g_df_immi <- g_df_immi %>% 
  dplyr::group_by(sim_id) %>% 
  dplyr::summarise(
    grand_mean_immi = base::mean(mean_immi),
    grand_sd_immi = fun_pooled_sd_equal_sample_sizes(sd_immi)
  )

g_df_pops <- g_df %>% 
  # dplyr::filter(year == 5 | year == 8) %>%
  dplyr::filter(year == 8) %>%
  dplyr::group_by(sim_id, study_num) %>% 
  dplyr::summarise(
    mean_popids = base::mean(pops_mean),
    sd_popids = stats::sd(pops_mean)
  )

g_df_pops <- g_df_pops %>% 
  dplyr::group_by(sim_id) %>% 
  dplyr::summarise(
    grand_mean_popids = base::mean(mean_popids),
    grand_sd_popids   = fun_pooled_sd_equal_sample_sizes(sd_popids)
  )

g_df_immi_perc_mean <- g_df_immi_perc %>% 
  dplyr::select(
    sim_id, 
    grand_mean_immi_perc, 
    grand_sd_immi_perc) %>% 
  dplyr::mutate(stat_measure = "immigrants")


colnames(g_df_immi_perc_mean)[colnames(g_df_immi_perc_mean)
%in% c("grand_mean_immi_perc", "grand_sd_immi_perc")] <- c("mean", "sd")

g_df_immi_perc_mean <- g_df_immi_perc_mean %>% 
  dplyr::mutate(
    is_within = ifelse(
      mean >= (g_immi_min * (100 - g_percentage) / 100) & 
        mean <= (g_immi_max * (100 + g_percentage) / 100),
      TRUE,
      FALSE
    )
  )

g_df_pops_mean <- g_df_pops %>% 
  dplyr::select(sim_id, grand_mean_popids, grand_sd_popids) %>% 
  dplyr::mutate(stat_measure = "populations")

colnames(g_df_pops_mean)[colnames(g_df_pops_mean)
%in% c("grand_mean_popids", "grand_sd_popids")] <- c("mean", "sd")

g_df_pops_mean <- g_df_pops_mean %>% 
  dplyr::mutate(
    is_within = ifelse(
      # mean >= 2 & mean <= (g_pop_max * (100 + g_percentage) / 100),
      mean >= 2,
      TRUE,
      FALSE
    )
  )

# read simulation parameters
g_files <- base::list.files(
  path = base::paste0(g_path, "/in/sim"),
  pattern = "*_simfile.csv",
  full.names = TRUE
)

g_params <- plyr::ldply(
  g_files,
  utils::read.table,
  sep = " ",
  fill = TRUE,
  header = TRUE
)

# take the parameter values for plotting
g_params <- g_params[, c("sim_id", "val1", "val2", "val3")]

# rename columns
base::colnames(g_params) <- c("sim_id", "mort", "birds_prop", "wind_prop")

# g_df <- base::merge(x = g_params, y = g_df, by = "sim_id")

# rm(g_df)

g_df_immi_abs <- base::merge(x = g_params, y = g_df_immi, by = "sim_id")
g_df_immi_abs$mort <- base::as.factor(g_df_immi_abs$mort)

g_df_cali <- base::rbind(g_df_immi_perc_mean, g_df_pops_mean)
g_df_cali <- base::merge(x = g_params, y = g_df_cali, by = "sim_id")
g_df_cali <- dplyr::arrange(g_df_cali, sim_id)
g_df_cali$stat_measure <- base::as.factor(g_df_cali$stat_measure)
g_df_cali$mort <- base::as.factor(g_df_cali$mort)
g_df_cali$birds_prop <- g_df_cali$birds_prop * 100
g_df_cali$wind_prop <- g_df_cali$wind_prop * 100


# - % immigration = [4.9%, 7.48%]
# - # pop ids per population: > 2 and < 4 pop ids per population
# - distance: mean = [900 m, 1210 m], 
#             sd = [666 m, 814 m], 
#             min = [288 m, 352 m], 
#             max = [2970 m, 3630 m]


########################################

cb_palette <- c(
  # "#999999",
  "#E69F00",
  "#56B4E9",
  "#009E73",
  "#F0E442",
  # "#0072B2",
  # "#D55E00",
  "#CC79A7"
)

g_df_cali_recode <- g_df_cali %>%
  dplyr::mutate(
    mort = dplyr::recode(
      mort, 
      "scn_cali_03_mort_0.csv" = "(a)", 
      "scn_cali_03_mort_2.csv" = "(b)",
      "scn_cali_03_mort_3.csv" = "(c)",
      "scn_cali_03_mort_4.csv" = "(d)"
    )
  )

p1 <- g_df_cali_recode %>% 
  dplyr::filter(is_within == TRUE) %>% 
  ggplot(mapping = aes(
    x = wind_prop, 
    y = birds_prop,
    col = stat_measure,
    color = stat_measure,
    shape = stat_measure
  )) +
  geom_point(size = 3) +
  # scale_fill_manual(
  #   # name = "Vector and location",
  #   #values =  c("#999999", "#E69F00", "#56B4E9", "#009E73")
  #   values = cb_palette
  # ) +
  scale_color_manual(
    name = "Measure:",
    # values =  c("#999999", "#E69F00", "#56B4E9", "#009E73")
    values = cb_palette
  ) +
  xlab("Seeds by wind (%)") +
  ylab("Cones by birds (%)") +
  coord_cartesian(xlim = c(0, 20), ylim = c(0, 20)) +
  # xlim(c(0, 0.2)) +
  # ylim(c(0, 0.2)) +
  # ggtitle("") +
  scale_shape_manual(
    name = "Measure:",
    values = c(4, 0)
  ) +
  theme_minimal() +
  scale_x_continuous(minor_breaks = seq(0, 20, 1)) +
  scale_y_continuous(minor_breaks = seq(0, 20, 1)) +
  theme(
    text         = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(
      color = "lightgrey", 
      size = 0.5, 
      linetype = 1),
    legend.position = "bottom"
  ) +
  ggplot2::facet_wrap(
    ~mort
  )

ggplot2::ggsave(
  plot = p1,
  filename = base::paste0(g_figure_name, "_dpi300.tiff"),
  path = base::paste0(g_path, "/plots"),
  width = 2 * 3,
  height = 6,
  bg = "white",
  dpi = 300
)



p2 <- g_df_cali %>% 
  dplyr::filter(stat_measure == "immigrants") %>%
  ggplot(mapping = aes(
    x = wind_prop, 
    y = birds_prop,
    fill = mean
  )) +
  geom_tile() +
  geom_text(aes(label = round(mean, 2)), color = "black", size = 4) +
  # scale_fill_gradient(low="white", high="blue") +
  viridis::scale_fill_viridis(discrete = FALSE) +
  xlab("proportion of seeds by wind") +
  ylab("proportion of cones by birds") +
  coord_cartesian(xlim = c(0, 20), ylim = c(0, 20)) +
  scale_shape_manual(
    name = "Measure",
    values = c(4, 0)
  ) +
  theme_minimal() +
  # hrbrthemes::theme_ipsum() +
  scale_x_continuous(minor_breaks = seq(0, 20, 1)) +
  scale_y_continuous(minor_breaks = seq(0, 20, 1)) +
  theme(
    text         = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(
      color = "lightgrey", 
      size = 0.5, 
      linetype = 1),
    legend.position = "bottom"
  ) +
  ggplot2::facet_wrap(
    ~mort
  )

p3 <- g_df_cali %>% 
  dplyr::filter(stat_measure == "populations") %>%
  ggplot(mapping = aes(
    x = wind_prop, 
    y = birds_prop,
    fill = mean
  )) +
  geom_tile() +
  # geom_text(aes(label = round(mean, 2)), color = "black", size = 4) +
  # scale_fill_gradient(low="white", high="blue") +
  viridis::scale_fill_viridis(discrete = FALSE) +
  xlab("proportion of seeds by wind") +
  ylab("proportion of cones by birds") +
  coord_cartesian(xlim = c(0, 20), ylim = c(0, 20)) +
  scale_shape_manual(
    name = "Measure",
    values = c(4, 0)
  ) +
  theme_minimal() +
  # hrbrthemes::theme_ipsum() +
  scale_x_continuous(minor_breaks = seq(0, 20, 1)) +
  scale_y_continuous(minor_breaks = seq(0, 20, 1)) +
  theme(
    text         = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(
      color = "lightgrey", 
      size = 0.5, 
      linetype = 1),
    legend.position = "bottom"
  ) +
  ggplot2::facet_wrap(
    ~mort
  )

####################################################

p4 <- g_df_immi_abs %>% 
  ggplot(mapping = aes(
    x = wind_prop, 
    y = birds_prop,
    fill = grand_mean_immi
  )) +
  geom_tile() +
  # geom_text(aes(label = round(mean, 2)), color = "black", size = 4) +
  # scale_fill_gradient(low="white", high="blue") +
  viridis::scale_fill_viridis(discrete = FALSE) +
  # scale_fill_brewer(palette = "Spectral") +
  # scale_fill_fermenter() +
  xlab("proportion of seeds by wind") +
  ylab("proportion of cones by birds") +
  # coord_cartesian(xlim = c(0, 20), ylim = c(0, 20)) +
  scale_shape_manual(
    name = "Measure",
    values = c(4, 0)
  ) +
  theme_minimal() +
  # hrbrthemes::theme_ipsum() +
  # scale_x_continuous(minor_breaks = seq(0, 20, 1)) +
  # scale_y_continuous(minor_breaks = seq(0, 20, 1)) +
  theme(
    text         = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(
      color = "lightgrey", 
      size = 0.5, 
      linetype = 1),
    legend.position = "bottom"
  ) +
  ggplot2::facet_wrap(
    ~mort
  ) 

g_df_immi_abs_mort_3 <- g_df_immi_abs %>% 
  dplyr::filter(mort == "scn_cali_03_mort_3.csv")

g_df_immi_abs_mort_4 <- g_df_immi_abs %>% 
  dplyr::filter(mort == "scn_cali_03_mort_4.csv")

# rename column
base::colnames(g_df_immi_abs_mort_4) <- c("sim_id_4", "mort_4", "birds_prop_4", "wind_prop_4", "grand_mean_4", "grand_sd_4")


g_df_immi_diff <- cbind(g_df_immi_abs_mort_3, g_df_immi_abs_mort_4)

g_df_immi_diff$mean_diff <- g_df_immi_diff$grand_mean_4 - g_df_immi_diff$grand_mean_immi

p4 <- g_df_immi_diff %>% 
  ggplot(mapping = aes(
    x = wind_prop, 
    y = birds_prop,
    fill = mean_diff
  )) +
  geom_tile() +
  # geom_text(aes(label = round(mean, 2)), color = "black", size = 4) +
  # scale_fill_gradient(low="white", high="blue") +
  # viridis::scale_fill_viridis(discrete = FALSE) +
   scale_fill_gradient2() +
  # scale_fill_brewer(palette = "Spectral") +
  # scale_fill_fermenter() +
  xlab("proportion of seeds by wind") +
  ylab("proportion of cones by birds") +
  # coord_cartesian(xlim = c(0, 20), ylim = c(0, 20)) +
  scale_shape_manual(
    name = "Measure",
    values = c(4, 0)
  ) +
  theme_minimal() +
  # hrbrthemes::theme_ipsum() +
  # scale_x_continuous(minor_breaks = seq(0, 20, 1)) +
  # scale_y_continuous(minor_breaks = seq(0, 20, 1)) +
  theme(
    text         = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(
      color = "lightgrey", 
      size = 0.5, 
      linetype = 1),
    legend.position = "bottom"
  )



##
```

```{r}
#| label: cali-3-plot
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(plyr)
library(dplyr)
# library(easycsv)
library(ggplot2)
library(cowplot)
library(hrbrthemes)
library(viridis)

# load utils
base::source("utils/pooled_sd.R")
base::source("utils/read_multiple_files.R")

# input files
g_path <- "data/sim_outputs/calibration/03_cali"
g_sim_type <- "*dispersal.csv"

# output figure
g_figure_name <- "03_cali_dispersal"

# read files and create df
g_df <- fun_read_files(paste0(g_path, "/raw"), g_sim_type)

g_df$immi <- g_df$bird_immi + g_df$wind_immi
g_df$lost <- g_df$wind_out + g_df$wind_unsui

# population size in 2006 (8-year-old plants) for each mortality scenario
g_df <- g_df %>% 
  dplyr::group_by(sim_id, study_num) %>% 
  dplyr::summarise(
    mean_immi = base::mean(immi),
    mean_lost = base::mean(lost)
)

g_df <- g_df %>% 
  dplyr::group_by(sim_id) %>% 
  dplyr::summarise(
    grand_mean_immi = base::mean(mean_immi),
    grand_mean_lost = base::mean(mean_lost)
  )

# read simulation parameters
g_files <- base::list.files(
  path = base::paste0(g_path, "/in/sim"),
  pattern = "*_simfile.csv",
  full.names = TRUE
)

g_params <- plyr::ldply(
  g_files,
  utils::read.table,
  sep = " ",
  fill = TRUE,
  header = TRUE
)

# take the parameter values for plotting
g_params <- g_params[, c("sim_id", "val1", "val2", "val3")]

# rename column
base::colnames(g_params) <- c("sim_id", "mort", "birds_prop", "wind_prop")

g_df <- base::merge(x = g_params, y = g_df, by = "sim_id")



# - % immigration = [4.9%, 7.48%]
# - # pop ids per population: > 2 and < 4 pop ids per population
# - distance: mean = [900 m, 1210 m], 
#             sd = [666 m, 814 m], 
#             min = [288 m, 352 m], 
#             max = [2970 m, 3630 m]


########################################

cb_palette <- c(
  # "#999999",
  "#E69F00",
  "#56B4E9",
  "#009E73",
  "#F0E442",
  # "#0072B2",
  # "#D55E00",
  "#CC79A7"
)


p1 <- g_df %>% 
  ggplot(mapping = aes(
    x = wind_prop, 
    y = birds_prop,
    fill = grand_mean_immi
  )) +
  geom_tile() +
  # geom_text(aes(label = round(mean, 2)), color = "black", size = 4) +
  # scale_fill_gradient(low="white", high="blue") +
  viridis::scale_fill_viridis(discrete = FALSE) +
  xlab("proportion of seeds by wind") +
  ylab("proportion of cones by birds") +
  # coord_cartesian(xlim = c(0, 20), ylim = c(0, 20)) +
  # scale_shape_manual(
    # name = "Measure",
    # values = c(4, 0)
  # ) +
  theme_minimal() +
  # hrbrthemes::theme_ipsum() +
  # scale_x_continuous(minor_breaks = seq(0, 20, 1)) +
  # scale_y_continuous(minor_breaks = seq(0, 20, 1)) +
  theme(
    text         = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(
      color = "lightgrey", 
      size = 0.5, 
      linetype = 1),
    legend.position = "bottom"
  ) +
  ggplot2::facet_wrap(
    ~mort
  )

p2 <- g_df %>% 
  ggplot(mapping = aes(
    x = wind_prop, 
    y = birds_prop,
    fill = grand_mean_lost
  )) +
  geom_tile() +
  # geom_text(aes(label = round(mean, 2)), color = "black", size = 4) +
  # scale_fill_gradient(low="white", high="blue") +
  viridis::scale_fill_viridis(discrete = FALSE) +
  xlab("proportion of seeds by wind") +
  ylab("proportion of cones by birds") +
  # coord_cartesian(xlim = c(0, 20), ylim = c(0, 20)) +
  # scale_shape_manual(
    # name = "Measure",
    # values = c(4, 0)
  # ) +
  theme_minimal() +
  # hrbrthemes::theme_ipsum() +
  # scale_x_continuous(minor_breaks = seq(0, 20, 1)) +
  # scale_y_continuous(minor_breaks = seq(0, 20, 1)) +
  theme(
    text         = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(
      color = "lightgrey", 
      size = 0.5, 
      linetype = 1),
    legend.position = "bottom"
  ) +
  ggplot2::facet_wrap(
    ~mort
  )


# ggplot2::ggsave(
#   plot = p1,
#   filename = base::paste0(g_figure_name, ".png"),
#   path = base::paste0(g_path, "/plots"),
#   width = 3 * 3,
#   height = 4,
#   bg = "white"
# )

##
```

# 4. Habitat quality

Calculate the direct proportion of different habitat quality considering
the mean flower count in Eneabba as the habitat quality with coefficient
1.

```{r}
#| label: habitat-quality
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
# library(plyr)
library(dplyr)
# library(easycsv)
library(ggplot2)
# library(cowplot)
# library(hrbrthemes)
# library(viridis)


folder_name <- "./data/field/"
file_name <- "Bankhook_FruityFireDB_09May2019.csv"

df_meta_flr <- utils::read.csv(base::paste0(folder_name, file_name))

# remove rows with NAs in columns "flr"
df_meta_flr <- dplyr::filter(df_meta_flr, flr != "NA")

graphics::hist(df_meta_flr$flr)

df_meta_flr %>%
ggplot2::ggplot() +
  ggplot2::geom_boxplot(aes(x = as.factor(visitYr), y = flr))

# remove the two outliers
df <- dplyr::filter(df_meta_flr, flr < 150)

df %>%
ggplot2::ggplot() +
  ggplot2::geom_boxplot(aes(x = as.factor(visitYr), y = flr))


df %>%
ggplot2::ggplot() +
  ggplot2::geom_boxplot(aes(x = as.factor(visitYr), y = flr)) +
  ggplot2::facet_wrap(~plotID)


df_2015 <- dplyr::filter(df, visitYr == 2015)
df_2016 <- dplyr::filter(df, visitYr == 2016)
df_2017 <- dplyr::filter(df, visitYr == 2017)


unique(df_2016$zone)

df_2015 %>% 
  dplyr::group_by(zone, plotID) %>% 
  dplyr::summarise(
    n_indiv = dplyr::n(),
    time_since_fire = base::max(TSF),
    elev = base::max(elevation),
    mean_flr = base::mean(flr),
    median_flr = stats::median(flr),
    q3 = stats::quantile(flr, 0.75)
  )

df_2016 %>% 
  dplyr::group_by(zone, plotID) %>% 
  dplyr::summarise(
    n_indiv = dplyr::n(),
    time_since_fire = base::max(TSF),
    elev = base::max(elevation),
    mean_flr = base::mean(flr),
    median_flr = stats::median(flr),
    q3 = stats::quantile(flr, 0.75)
  )

df_2017 %>% 
  dplyr::group_by(zone, plotID) %>% 
  dplyr::summarise(
    n_indiv = dplyr::n(),
    time_since_fire = base::max(TSF),
    elev = base::max(elevation),
    mean_flr = base::mean(flr),
    median_flr = stats::median(flr),
    q3 = stats::quantile(flr, 0.75)
  )

df_2016_eneabba <- dplyr::filter(
  df_2016,
  TSF > 10 & 
    (zone == "Metapop" | 
       zone == "MetapopSouth" | 
       zone == "SouthEneabbaReserve"
    )
)

df_2016_eneabba %>% 
  dplyr::group_by(zone, plotID) %>% 
  dplyr::summarise(
    n_indiv = dplyr::n(),
    time_since_fire = base::max(TSF),
    elev = base::max(elevation),
    mean_flr = base::mean(flr),
    median_flr = stats::median(flr),
    q3 = stats::quantile(flr, 0.75)
  )

p1 <- df_2016_eneabba %>%
  ggplot() +
  geom_boxplot(aes(x = plotID, y = flr, color = zone), outlier.shape = NA) +
  geom_jitter(aes(x = plotID, y = flr), size = 1, alpha = 0.5) +
  scale_color_manual(
    name = "Location",
    labels = c("Eneabba", "South Eneabba", "Eneabba Reserve"),
    values = c("#868686FF", "#0073C2FF", "#EFC000FF")
    # values = c("#56B4E9", "#E69F00")
  ) +
  scale_y_continuous(
    breaks = base::seq(from = 0, to = 50, by = 10)
  ) +
  xlab("Plot ID") +
  ylab("# of flowers") +
  ggplot2::theme_classic() +
  ggplot2::theme(
    text = element_text(size = 9, face = "bold"),
    axis.text = element_text(size = 9),
    axis.title = element_text(size = 9, face = "bold"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9, face = "bold"),
    strip.text.x = element_blank(),
    strip.background = element_rect(colour = "white", fill = "white"),
    legend.position = c(0.75, 0.8)
  )

p2 <- df_2016_eneabba %>%
  ggplot() +
  geom_boxplot(aes(x = as.factor(elevation), y = flr)) +
  geom_jitter(aes(x = as.factor(elevation), y = flr))

## mean flowers of low habitat quality
x1 <- mean(df_2016_eneabba$flr[df_2016_eneabba$plotID == "HD1"])

## mean flowers of moderate habitat quality
x2 <- mean(df_2016_eneabba$flr[df_2016_eneabba$elevation > 60])

## mean flowers of high habitat quality
x3 <- mean(df_2016_eneabba$flr[df_2016_eneabba$plotID == "MPS1993"])

# calculate the habitat quality coefficients 
g_coef_moderate <- x2 / x1
g_coef_high <- x3 / x1

p1

base::print(
  base::paste0(
    "'Moderate' habitat quality coefficient is: ", 
    round(g_coef_moderate, 2)
  )
)
base::print(
  base::paste0(
    "'High' habitat quality coefficient is: ",
    round(g_coef_high, 2)
  )
)

ggplot2::ggsave(
  plot = p1,
  filename = base::paste0("habitat_quality_small_dpi600.tiff"),
  path = "./data/plots",
  width = 3.31,
  height = 3.31,
  bg = "white",
  dpi = 600
)

```

# 5. Flower count distributions

```{r}
#| label: load-clean-flower-data
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)
# library(easycsv)
library(ggplot2)
# library(cowplot)
# library(hrbrthemes)
# library(viridis)

# load daily rainfall climate data
# Eneabba Weather Station no. 8225
rain <- utils::read.csv(
  file = "./data/field/IDCJAC0009_8225_1800_Data.csv", 
  header = TRUE, 
  sep = ","
)

flower_current <- utils::read.csv(
  file = "./data/field/bankhook_Eneabba_flowers_2008_2020.csv", 
  header = TRUE, 
  sep = ","
)

# rename column names from rainfall data
rain <- rain[c(3, 4, 5, 6)]
colnames(rain)[1] <- "year"
colnames(rain)[2] <- "month"
colnames(rain)[3] <- "day"
colnames(rain)[4] <- "rainfall"

# tidy rainfall data --------------------------------------
rain <- rain %>%
  dplyr::group_by(year, month) %>%
  dplyr::summarise(rainfall = base::sum(rainfall, na.rm = TRUE))

# winterplus is winter-spring period taking one month before winter (i.e. May)
# and one month after spring (i.e. December) to get earlier and later rainfalls.
# This was a?ready used for the baseline flower data in Keith et al., (2014) 
rain_winterplus <- subset(rain, month > 4)

rain_winterplus <- rain_winterplus %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(winterplus_sum = base::sum(rainfall, na.rm = TRUE))

rain_annual <- rain %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(annual_sum = base::sum(rainfall, na.rm = TRUE))

rain <- base::merge(
  x = rain_winterplus,
  y = rain_annual,
  by = "year",
  all.y = TRUE
)

rain <- base::subset(rain, year < 2017)

base::remove(rain_annual, rain_winterplus)

# I sum up +1 year to the column Year. This makes annual and winterplus to be lag values (i.e., previous year)
rain_lag <- dplyr::mutate(rain, year = year + 1)
colnames(rain_lag)[2] <- "lag_winterplus_sum"
colnames(rain_lag)[3] <- "lag_annual_sum"

rain_lag <- rain_lag %>% 
  dplyr::mutate(
    lag2_annual_sum = stats::lag(lag_annual_sum),
    lag3_annual_sum = stats::lag(lag2_annual_sum),
    ac_annual_lag12 = lag_annual_sum + lag2_annual_sum,
    ac_annual_lag23 = lag2_annual_sum + lag3_annual_sum,
    ac_annual_lag123 = lag_annual_sum + lag2_annual_sum + lag3_annual_sum,
    lag2_winterplus_sum = stats::lag(lag_winterplus_sum),
    lag3_winterplus_sum = stats::lag(lag2_winterplus_sum),
    ac_winterplus_lag12 = lag_winterplus_sum + lag2_winterplus_sum,
    ac_winterplus_lag23 = lag2_winterplus_sum + lag3_winterplus_sum,
    ac_winterplus_lag123 = lag_winterplus_sum + lag2_winterplus_sum + 
      lag3_winterplus_sum
  ) %>% 
  stats::na.omit()

rain <- base::merge(
  x = rain,
  y = rain_lag,
  by = "year",
  all = TRUE
)

base::remove(rain_lag)


# tidy flower data ----------------------------------------
flower_current_all <- flower_current
flower_current_all$year_num <- flower_current_all$year
flower_current_all$year <- base::as.factor(flower_current_all$year) 
flower_current$year <- base::as.factor(flower_current$year)

flower_current <- flower_current[stats::complete.cases(
  flower_current[ , "flowers"]), ]

flower_current <- base::merge(
  x  = flower_current,
  y  = rain,
  by = "year"
)

# # The plants 604, 606, and 608 are outliers. These three plants are "rain on".
# # This means these plants were watered every month from other study.
# boxplot(flowers ~ year, data = flower_current,
#         xlab="Year", ylab="Number of flowers")


# # The plants 604, 606, and 608 are outliers. These three plants are "rain on".
# # This means these plants were watered every month from other study.
flower_current <- base::subset(
  flower_current, 
  tag != 604 & tag != 606 & tag != 608
)

flower_current$tag <- base::as.factor(flower_current$tag)

flower_current_all <- base::subset(
  flower_current_all, 
  tag != 604 & tag != 606 & tag != 608
)

flower_current_all$tag <- base::as.factor(flower_current_all$tag)

## two plants were recorded for first time already dead
## they are 893 and 894
flower_current_all <- flower_current_all %>% 
  dplyr::filter(!(tag %in% c("893", "894")))

## and three new plants in 2018 did not include the number of flowers
## they had zero flowers in 2018, thus it should be zero in order to 
## make the calculations, and avoid NAs and infinite values in later calculations
## plants 605.18, 606.18, and 614.18

## the plant tag "9771" in HD1 started in 2018 and and in 2019 the 
## classification status is again "new". I remove the "new" in the input file
flower_current_all <- flower_current_all %>% 
  dplyr::mutate(
    flowers = base::ifelse(
      status == "" & base::is.na(flowers),
      0,
      base::ifelse(
        status == "new" & base::is.na(flowers),
        0,
        flowers
      )
    )
  )

```

```{r}
#| label: flower-producer-class
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)
# library(easycsv)
library(ggplot2)
# library(cowplot)
# library(hrbrthemes)
# library(viridis)

# load and clean data
# source("./scripts/clean_data/01_import.R")
# source("./scripts/clean_data/02_clean_names.R")
# source("./scripts/clean_data/03_tidy.R")

df_all <- flower_current_all %>% 
  dplyr::mutate(cones = base::ifelse(
    status == "new" | status == "prevtagged not in database",
    flowers + TotFertile,
    flowers
  )
)

# remove DNF rows "not found tags", or "data not found"
df_all <- dplyr::filter(df_all, status != "DNF")

# only tags from 2008, i.e. initial plants
df_tags_only_2008 <- flower_current_all %>% 
  dplyr::filter(flower_current_all$year == 2008)

# cone accumulation per individual (i.e. tag)
df_all <- df_all %>% 
  dplyr::group_by(tag) %>% 
  dplyr::mutate(
    cum_flowers = base::cumsum(flowers),
    cum_cones = base::cumsum(cones)
  )

df_all_2017 <- dplyr::filter(df_all, year_num <= 2017)

# survival per individuals
## all plants in HD1 all years 2008-2020
# df_all <- df_all %>% 
#   dplyr::mutate(
#     survival = base::ifelse(
#       status == "",
#       1,
#       base::ifelse(
#         status == "dead", 
#         (year_num - base::min(year_num)) / 
#           (base::max(year_num) - base::min(year_num) + 1),
#         base::ifelse(
#           status == "new" 
#           | status == "fallen" 
#           | status == "prevtagged not in database",
#           1,
#           base::ifelse(
#             status == "dead2019" & year_num == 2019,
#             (year_num - base::min(year_num)) / 
#               (base::max(year_num) - base::min(year_num) + 1),
#             base::ifelse(
#               status == "dead2020" & year_num == 2020,
#               (year_num - base::min(year_num)) / 
#                 (base::max(year_num) - base::min(year_num) + 1),
#               NA
#             )
#           )
#         )
#       )
#     )
#   )

# survival per individuals
## all plants in HD1, from 2008 to 2017
df_all_2017 <- df_all_2017 %>% 
  dplyr::mutate(survival = base::ifelse(
    status == "",
    1,
    base::ifelse(
      status == "dead", 
      (year_num - base::min(year_num)) / 
        (base::max(year_num) - base::min(year_num) + 1),
      base::ifelse(
        status == "new" | status == "fallen",
        1,
        NA
        )
      )
    )
  )

# cumulative flowers only, and cumulative cones and flowers
# minimum survival in order to get the the survival rate when the plant died

## all plants, from 2008 to 2017
df_summ_all_2017 <- df_all_2017 %>% 
  dplyr::group_by(tag) %>% 
  dplyr::summarise(
    tag_cum_flowers = base::max(cum_flowers, na.rm = TRUE),
    tag_cum_cones = base::max(cum_cones, na.rm = TRUE),
    tag_height = base::max(hgt, na.rm = TRUE),
    tag_survival = base::min(survival, na.rm = TRUE)
  )

# quantile classification
df_q_all_2017 <- df_summ_all_2017 %>%
  dplyr::summarise(
    cones_q3 = stats::quantile(tag_cum_cones, prob = 0.75),
    cones_q1 = stats::quantile(tag_cum_cones, prob = 0.25),
    flowers_q3 = stats::quantile(tag_cum_flowers, prob = 0.75),
    flowers_q1 = stats::quantile(tag_cum_flowers, prob = 0.25)
  )

df_summ_all_2017 <- df_summ_all_2017 %>% 
  dplyr::mutate(
    q_class_cones = base::ifelse(
      tag_cum_cones > df_q_all_2017$cones_q3,
      3,
      base::ifelse(
        tag_cum_cones < df_q_all_2017$cones_q1,
        1,
        2
      )
    )
  ) %>% 
  dplyr::mutate(
    q_class_flowers = base::ifelse(
      tag_cum_flowers > df_q_all_2017$flowers_q3,
      3,
      base::ifelse(
        tag_cum_flowers < df_q_all_2017$flowers_q1,
        1,
        2
      )
    )
  )

# quantile classification
df_summ_all_2017 <- df_summ_all_2017 %>% 
  dplyr::mutate(
    q75_class_cones = base::ifelse(
      tag_cum_cones < df_q_all_2017$cones_q3,
      0,
      1
    )
  )

# plots with quantile classification
df_summ_all_2017 %>%
  dplyr::group_by(q_class_cones) %>%
  dplyr::summarise(
    total_cases = dplyr::n(),
    mean_survival = base::mean(tag_survival)
  )

# plot
p1 <- df_summ_all_2017 %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = tag_cum_cones, y = tag_survival)
  ) +
  ggplot2::geom_point(
    ggplot2::aes(color = base::as.factor(q75_class_cones)), 
    size = 2,
    alpha = 0.5
  ) +
  # ggrepel::geom_text_repel(
  #   mapping = ggplot2::aes(label = tag)
  # ) +
  ggplot2::coord_cartesian(ylim = c(0, 1)) +
  ggplot2::ggtitle("(a)") +
  ggplot2::xlab("Total cones (#)") +
  ggplot2::ylab("Survival rate") +
  ggplot2::theme_classic() +
  ggplot2::theme(
    text         = element_text(size = 14, face = "bold"),
    axis.text    = element_text(size = 14),
    axis.title   = element_text(size = 14, face = "bold"),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
    legend.position = c(0.75, 0.3)
    # legend.key.size = unit(1, "cm"),
    # legend.text  = element_text(size = 24, face = "plain"),
    # legend.title = element_text(size = 24, face = "bold")
  ) +
  # ggplot2::scale_color_manual(
  #   name = "Percentile class:",
  #   values = c("#0073C2FF", "#EFC000FF"),
  #   labels = c(base::expression("" < 0.75), base::expression("" >= 0.75))
  # )
    ggplot2::scale_color_manual(
    name = "Flower producer class:",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("poor", "good")
  )



# ggplot2::ggsave(
#   plot = p1,
#   filename = base::paste0(g_figure_name, ".png"),
#   path = base::paste0(g_path, "/plots"),
#   width = 7,
#   height = 4,
#   bg = "white"
# )

##
```

```{r}
#| label: flower-producer-class-2
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)
library(ggplot2)

flower_current_all_rain <- flower_current_all[
  stats::complete.cases(flower_current_all[ , "flowers"]), 
]

flower_current_all_rain <- base::merge(
  x  = flower_current_all_rain,
  y  = rain,
  by = "year"
)

flower_current_q_all_2017 <- base::merge(
  x = flower_current_all_rain, 
  y = df_summ_all_2017,
  by = "tag")


p2 <- flower_current_q_all_2017 %>% 
  ggplot(aes(x = year, y = flowers)) +
  geom_jitter(
    aes(color = as.factor(q75_class_cones)), width = 0.1, size = 2, alpha = 0.5
  ) +
  ggplot2::ggtitle("(b)") +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Flowers (#)") +
  scale_x_discrete(labels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)) +
  # ggplot2::theme_minimal() +
  ggplot2::theme_classic() +
  ggplot2::theme(
    text         = element_text(size = 14, face = "bold"),
    axis.text    = element_text(size = 14),
    axis.title   = element_text(size = 14, face = "bold"),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
  ) +
  # ggplot2::scale_color_manual(
  #   name = "Percentile class:",
  #   values = c("#0073C2FF", "#EFC000FF"),
  #   labels = c(base::expression("" < 0.75), base::expression("" >= 0.75))
  # )
    ggplot2::scale_color_manual(
    name = "Flower producer class:",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("poor", "good")
  )



```

```{r}
#| label: flower-data-boxplot
#| include: true
#| echo: false
#| collapse: false
#| warning: false

library(ggsignif)
library(ggplot2)
p3 <- ggplot(data = df_summ_all_2017,
       aes(x = as.factor(q75_class_cones),
           y = tag_cum_cones,
           fill = as.factor(q75_class_cones))) +
  geom_boxplot(
    alpha = 0.6,
    outlier.shape = 1,
    notch = TRUE,
    notchwidth = 0.5
  ) +
  #  geom_jitter(
  #   width = 0.2,
  #   size = 1,
  #   alpha = 0.7,
  #   shape = 21,
  #   color = "black",
  #   fill = "black"
  # ) + 
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF")) +
  guides(fill = FALSE) +
  theme_classic() +
  ggtitle("(c)") +
  xlab("Flower producers class") +
  ylab("Total cones (#)") +
  scale_x_discrete(labels = c("poor", "good")) +
  theme(text = element_text(size = 14, face = "bold"),
        axis.text    = element_text(size = 14),
        axis.title   = element_text(size = 14, face = "bold"),
        legend.text  = element_text(size = 14),
        legend.title = element_text(size = 14, face = "bold")) +
  ggsignif::geom_signif(test = "wilcox.test", comparisons = list(c(1, 2)), 
              map_signif_level = TRUE, y_position = 92)


p4 <- ggplot(data = df_summ_all_2017,
       aes(x = as.factor(q75_class_cones),
           y = tag_survival,
           fill = as.factor(q75_class_cones))) +
  geom_boxplot(
    alpha = 0.6,
    outlier.shape = 1,
    notch = TRUE,
    notchwidth = 0.5
  ) +
  #  geom_jitter(
  #   width = 0.2,
  #   size = 1,
  #   alpha = 0.7,
  #   shape = 21,
  #   color = "black",
  #   fill = "black"
  # ) + 
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF")) +
  guides(fill = FALSE) +
  theme_classic() +
  ggtitle("(d)") +
  xlab("Flower producers class") +
  ylab("Survival rate") +
  scale_x_discrete(labels = c("poor", "good")) +
  ggplot2::scale_y_continuous(
    breaks = base::seq(from = 0, to = 1, by = 0.25)
  ) +
  coord_cartesian(ylim = c(0, 1.1)) +
  theme(text = element_text(size = 14, face = "bold"),
        axis.text    = element_text(size = 14),
        axis.title   = element_text(size = 14, face = "bold"),
        legend.text  = element_text(size = 14),
        legend.title = element_text(size = 14, face = "bold")) +
  ggsignif::geom_signif(test = "wilcox.test", comparisons = list(c(1, 2)), 
              map_signif_level = TRUE, y_position = 1.05)
```

```{r}
#| label: flower-data-plot
#| include: true
#| echo: false
#| collapse: false
#| warning: false

library(ggpubr)

parr1 <- ggpubr::ggarrange(p1, p2, ncol = 2, common.legend = TRUE, legend = "bottom")
parr2 <- ggpubr::ggarrange(p3, p4, ncol = 2, common.legend = TRUE, legend = "bottom")
parr3 <- ggpubr::ggarrange(parr1, parr2, ncol = 1)

# save figure
ggplot2::ggsave(
  plot = parr3,
  filename = base::paste0("flower_producers_NEW_dpi300", ".tiff"),
  path = "./data/plots/",
  width = 3.31 * 2,
  height = 3.31 * 2,
  bg = "white",
  dpi = 300
)

```

```{r}
#| label: flower-data-glm-fit
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# libraries
library(rstatix)

# Testing normality: both datasets are significant, thus they are non-normal distributed.
stats::shapiro.test(flower_current_q_all_2017$flowers)
stats::shapiro.test(flower_current_q_all_2017$tag_cum_cones)
stats::shapiro.test(flower_current_q_all_2017$tag_survival)

# points are not along the reference line in que qqplot
car::qqPlot(flower_current_q_all_2017$flowers)
car::qqPlot(flower_current_q_all_2017$tag_cum_cones)

res1 <- stats::wilcox.test(
  tag_cum_cones ~ q75_class_cones, 
  data = flower_current_q_all_2017,
  exact = FALSE
)

res2 <- stats::wilcox.test(
  tag_survival ~ q75_class_cones, 
  data = flower_current_q_all_2017,
  exact = FALSE
)

stat_test1 <- flower_current_q_all_2017 %>% 
  rstatix::wilcox_test(tag_cum_cones ~ q75_class_cones) %>% 
  rstatix::add_significance()

stat_test2 <- flower_current_q_all_2017 %>% 
  rstatix::wilcox_test(tag_survival ~ q75_class_cones) %>% 
  rstatix::add_significance()

stats::IQR(flower_current_q_all_2017$tag_cum_cones)
stats::IQR(flower_current_q_all_2017$tag_survival)

# generalize linear regression with the binomial family
# shows the predictor variable year as non significant, thus,
# we can assume that the higher the flower production the higher the survival
# probability it is with year being not significant in the binary classification
# year variable does not influence in the binary calssification of lower and 
# higher flower producers 
fit_glm <- stats::glm(
  q75_class_cones ~ flowers + year_num,
  family = binomial,
  data = flower_current_q_all_2017
)

summary(fit_glm)
```

## 5.1 Fuzzification

```{r}
#| label: fuzzification
#| include: true
#| echo: false
#| collapse: false
#| warning: false

library(ggplot2)
library(dplyr)

p1 <- flower_current_q_all_2017 %>% 
  ggplot(aes(x = ac_winterplus_lag123, y = flowers)) +
  geom_jitter(
    aes(color = as.factor(q75_class_cones)), width = 0.2, size = 2, alpha = 0.5
  ) +
  ggplot2::ggtitle("(a)") +
  ggplot2::xlab("Sum winter\u2013spring rainfall\n of the last three years (mm)") +
  ggplot2::ylab("Flowers (#)") +
  # ggplot2::theme_minimal() +
  ggplot2::theme_classic() +
  ggplot2::theme(
    text         = element_text(size = 14, face = "bold"),
    axis.text    = element_text(size = 14),
    axis.title   = element_text(size = 14, face = "bold"),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
  ) +
  # ggplot2::scale_color_manual(
  #   name = "Percentile class:",
  #   values = c("#0073C2FF", "#EFC000FF"),
  #   labels = c(base::expression("" < 0.75), base::expression("" >= 0.75))
  # )
  ggplot2::scale_color_manual(
    name = "Flower producer",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("low", "high")
  )

# classify fuzzy sets
# >=1134.9 mm
flower_current_q_all_2017 <- flower_current_q_all_2017 %>% 
  dplyr::mutate(fuzzy = base::ifelse(ac_winterplus_lag123 >= 1134.9, 1, 0))

p2 <- flower_current_q_all_2017 %>%
  dplyr::filter(fuzzy == 0) %>% 
  ggplot(aes(x = lag_annual_sum, y = flowers)) +
  geom_jitter(
    aes(color = as.factor(q75_class_cones)), width = 0.2, size = 2, alpha = 0.5
  ) +
  ggplot2::ggtitle("(b)") +
  ggplot2::xlab("Lag annual rainfall (mm)") +
  ggplot2::ylab("Flowers (#)") +
  # ggplot2::theme_minimal() +
  ggplot2::theme_classic() +
  ggplot2::theme(
    text         = element_text(size = 14, face = "bold"),
    axis.text    = element_text(size = 14),
    axis.title   = element_text(size = 14, face = "bold"),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
  ) +
  # ggplot2::scale_color_manual(
  #   name = "Percentile class:",
  #   values = c("#0073C2FF", "#EFC000FF"),
  #   labels = c(base::expression("" < 0.75), base::expression("" >= 0.75))
  # )
  ggplot2::scale_color_manual(
    name = "Flower producer",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("low", "high")
  )

p3 <- flower_current_q_all_2017 %>%
  dplyr::filter(fuzzy == 1) %>% 
  ggplot(aes(x = lag_annual_sum, y = flowers)) +
  geom_jitter(
    aes(color = as.factor(q75_class_cones)), width = 0.2, size = 2, alpha = 0.5
  ) +
  ggplot2::ggtitle("(c)") +
  ggplot2::xlab("Lag annual rainfall (mm)") +
  ggplot2::ylab("Flowers (#)") +
  # ggplot2::theme_minimal() +
  ggplot2::theme_classic() +
  ggplot2::theme(
    text         = element_text(size = 14, face = "bold"),
    axis.text    = element_text(size = 14),
    axis.title   = element_text(size = 14, face = "bold"),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
  ) +
  # ggplot2::scale_color_manual(
  #   name = "Percentile class:",
  #   values = c("#0073C2FF", "#EFC000FF"),
  #   labels = c(base::expression("" < 0.75), base::expression("" >= 0.75))
  # )
  ggplot2::scale_color_manual(
    name = "Flower producers",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("low", "high")
  )


parr <- ggpubr::ggarrange(
  p1, p2, p3,
  nrow = 3, 
  common.legend = TRUE,
  legend = "bottom"
)

# save figure
# ggplot2::ggsave(
#   plot = parr,
#   filename = base::paste0("fuzzification_small", ".png"),
#   path = "./data/plots/",
#   width = 3.31,
#   height = 3.31 * 3,
#   bg = "white"
# )

```

```{r}
#| label: fuzzification-plot
#| include: true
#| echo: false
#| collapse: false
#| warning: false


p1f <- flower_current_q_all_2017 %>% 
  ggplot(aes(x = ac_winterplus_lag123, y = flowers)) +
    ggplot2::geom_vline(
    mapping = aes(
      xintercept = 1134.9, 
      linetype = "crisp values"), 
    colour = "black") +
      ggplot2::geom_vline(
    mapping = aes(
      xintercept = 1101.6, 
      linetype = "crisp values"), 
    colour = "black") +
  geom_jitter(
    aes(color = as.factor(q75_class_cones)), width = 0.3, size = 3, alpha = 0.5
  ) +
    scale_y_continuous(
    "Flowers (#)", 
    sec.axis = sec_axis(~ . /22, name = "Proportion")
  ) +
  ggplot2::ggtitle("(a)") +
  ggplot2::xlab("Sum winter\u2013spring rainfall of the last three years (mm)") +
  ggplot2::ylab("Flowers (#)") +
  # ggplot2::theme_minimal() +
  ggplot2::theme_bw() +
  ggplot2::theme(
    text         = element_text(size = 14, face = "bold"),
    axis.text    = element_text(size = 14),
    axis.title   = element_text(size = 14, face = "bold"),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
  ) +
  # ggplot2::scale_color_manual(
  #   name = "Percentile class:",
  #   values = c("#0073C2FF", "#EFC000FF"),
  #   labels = c(base::expression("" < 0.75), base::expression("" >= 0.75))
  # )
  ggplot2::scale_color_manual(
    name = "Flower producer:",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("poor", "good")
  ) +
  ggplot2::scale_linetype_manual(name = "", values = c("dashed")) 

# classify fuzzy sets
# >=1134.9 mm
flower_current_q_all_2017 <- flower_current_q_all_2017 %>% 
  dplyr::mutate(fuzzy = base::ifelse(ac_winterplus_lag123 >= 1134.9, 1, 0))

p2f <- flower_current_q_all_2017 %>%
  dplyr::filter(fuzzy == 0) %>% 
  ggplot(aes(x = lag_annual_sum, y = flowers)) +
  ggplot2::geom_vline(
    mapping = aes(
      xintercept = 341.5, 
      linetype = "crisp values"), 
    colour = "black") +
  ggplot2::geom_vline(
    mapping = aes(
      xintercept = 369.3, 
      linetype = "crisp values"), 
    colour = "black") +
  ggplot2::geom_vline(
    mapping = aes(
      xintercept = 480.3, 
      linetype = "crisp values"), 
    colour = "black") +
  geom_jitter(
    aes(color = as.factor(q75_class_cones)), width = 0.3, size = 3, alpha = 0.5
  ) +
  # ggplot2::xlim(c(340, 500)) +
  scale_y_continuous(
    "Flowers (#)", 
    sec.axis = sec_axis(~ . / 7, name = "Proportion")
  ) +
  ggplot2::scale_x_continuous(breaks = base::seq(325, 525, 50)) +
  ggplot2::ggtitle("(b)") +
  ggplot2::xlab("Lag annual rainfall (mm)") +
  ggplot2::ylab("Flowers (#)") +
  ggplot2::theme_bw() +
  ggplot2::theme(
    text         = element_text(size = 14, face = "bold"),
    axis.text    = element_text(size = 14),
    axis.title   = element_text(size = 14, face = "bold"),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
  ) +
  ggplot2::scale_color_manual(
    name = "Flower producer:",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("poor", "good") 
  ) +
  ggplot2::scale_linetype_manual(name = "", values = c("dashed")) +
  ggplot2::coord_cartesian(xlim = c(325, 525))

p3f <- flower_current_q_all_2017 %>%
  dplyr::filter(fuzzy == 1) %>% 
  ggplot(aes(x = lag_annual_sum, y = flowers)) +
  ggplot2::geom_vline(
    mapping = aes(
      xintercept = 412.8, 
      linetype = "crisp values"), 
    colour = "black") +
  ggplot2::geom_vline(
    mapping = aes(
      xintercept = 458.3, 
      linetype = "crisp values"), 
    colour = "black") +
  ggplot2::geom_vline(
    mapping = aes(
      xintercept = 492.7, 
      linetype = "crisp values"), 
    colour = "black") +
  ggplot2::geom_vline(
    mapping = aes(
      xintercept = 518.3, 
      linetype = "crisp values"), 
    colour = "black") +
  geom_jitter(
    aes(color = as.factor(q75_class_cones)), width = 0.3, size = 3, alpha = 0.5
  ) +
  scale_y_continuous(
    "Flowers (#)", 
    sec.axis = sec_axis(~ . /22, name = "Proportion")
  ) +
  ggplot2::scale_x_continuous(breaks = base::seq(325, 525, 50)) +
  # ggplot2::xlim(c(385, 530)) +
  ggplot2::ggtitle("(c)") +
  ggplot2::xlab("Lag annual rainfall (mm)") +
  ggplot2::ylab("Flowers (#)") +
  # ggplot2::theme_minimal() +
  ggplot2::theme_bw() +
  ggplot2::theme(
    text         = element_text(size = 14, face = "bold"),
    axis.text    = element_text(size = 14),
    axis.title   = element_text(size = 14, face = "bold"),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
  ) +
  # ggplot2::scale_color_manual(
  #   name = "Percentile class:",
  #   values = c("#0073C2FF", "#EFC000FF"),
  #   labels = c(base::expression("" < 0.75), base::expression("" >= 0.75))
  # )
  ggplot2::scale_color_manual(
    name = "Flower producer:",
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("poor", "good")
  ) +
  ggplot2::scale_linetype_manual(name = "", values = c("dashed")) +
  ggplot2::coord_cartesian(xlim = c(375, 525))


parr <- ggpubr::ggarrange(
  p1f, p2f, p3f,
  nrow = 3, 
  common.legend = TRUE,
  legend = "bottom"
)

# save figure
ggplot2::ggsave(
  plot = parr,
  filename = base::paste0("fuzzification_all_dpi300", ".tiff"),
  path = "./data/plots/",
  width = 3.31 * 2,
  height = 3.31 * 3,
  bg = "white",
  dpi = 300
)

```

```{r}
#| label: fuzzification-plot-fuzzyr
#| include: false
#| echo: false
#| collapse: false
#| warning: false

# library(FuzzyR)
# 
# g_fis <- FuzzyR::newfis("tipper")
# g_fis <- FuzzyR::addvar(
#   fis = g_fis, 
#   varType = "input", 
#   varName = "rainfall", 
#   varBounds = 300:550
# )
# 
# g_fis <- FuzzyR::addmf(g_fis, "input", 1, "poor", "trapmf", c(300, 350, 400, 450))
# g_fis <- FuzzyR::addmf(g_fis, "input", 1, "good", "trapmf", c(400, 450, 500, 550))
# 
# class(g_fis)
# 
# FuzzyR::plotmf(g_fis, "input", 1)
# 
# plotmf(
# g_fis,
# "input",
# 1,
# xx = 0,
# timelimit = 0,
# xlab = "hey",
# ylab = NULL,
# main = NULL
# )
```

## 5.2 Fit flower count distributions

### 5.1.1 fs 0, mf 0

Null, no flowers in membership function zero for both plant types

```{r}
#| label: fs0-mf0
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 0, membership function = 0, plant type  = 0

# below 341.5 mm
df_dist_fs0_mf0 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 0 & lag_annual_sum == 341.5
  )

# no flowers in this membership function
base::max(df_dist_fs0_mf0$flowers)

```

### 5.1.2 fs 0, mf 1

```{r}
#| label: fs0-mf1-pt0
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 0, membership function = 1, plant type  = 0
# df_dist <- df_dist %>% 
#   dplyr::mutate(
#     class_lag123 = base::ifelse(
#       ac_winterplus_lag123 > 1134,
#       1,
#       0
#     )
#   )

# lag_annual_sum = 369.3 mm
df_dist_fs0_mf1 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 0 & lag_annual_sum == 369.3
    )

fs0_mf1_pt0_p <- fitdistrplus::fitdist(
  df_dist_fs0_mf1$flowers[df_dist_fs0_mf1$q75_class_cones == 0],
  "pois"
)

fs0_mf1_pt0_nb <- fitdistrplus::fitdist(
  df_dist_fs0_mf1$flowers[df_dist_fs0_mf1$q75_class_cones == 0],
  "nbinom"
)

fs0_mf1_pt0_g <- fitdistrplus::fitdist(
  df_dist_fs0_mf1$flowers[df_dist_fs0_mf1$q75_class_cones == 0],
  "geom"
)

plot(fs0_mf1_pt0_p)
plot(fs0_mf1_pt0_nb)
plot(fs0_mf1_pt0_g)

summary(fs0_mf1_pt0_p)
summary(fs0_mf1_pt0_nb)
summary(fs0_mf1_pt0_g)

# get aic values for each distribution
fs0_mf1_pt0_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs0_mf1_pt0_p$aic
)
fs0_mf1_pt0_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs0_mf1_pt0_nb$aic
)
fs0_mf1_pt0_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs0_mf1_pt0_g$aic
)


g_aic <- base::rbind(
  fs0_mf1_pt0_p_aic, 
  fs0_mf1_pt0_nb_aic, 
  fs0_mf1_pt0_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

```{r}
#| label: fs0-mf1-pt1
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 0, membership function = 1, plant type  = 1
df_dist_fs0_mf1 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 0 & lag_annual_sum == 369.3
    )

fs0_mf1_pt1_p <- fitdistrplus::fitdist(
  df_dist_fs0_mf1$flowers[df_dist_fs0_mf1$q75_class_cones == 1],
  "pois"
)

fs0_mf1_pt1_nb <- fitdistrplus::fitdist(
  df_dist_fs0_mf1$flowers[df_dist_fs0_mf1$q75_class_cones == 1],
  "nbinom"
)

fs0_mf1_pt1_g <- fitdistrplus::fitdist(
  df_dist_fs0_mf1$flowers[df_dist_fs0_mf1$q75_class_cones == 1],
  "geom"
)

plot(fs0_mf1_pt1_p)
plot(fs0_mf1_pt1_nb)
plot(fs0_mf1_pt1_g)

summary(fs0_mf1_pt1_p)
summary(fs0_mf1_pt1_nb)
summary(fs0_mf1_pt1_g)

# get aic values for each distribution
fs0_mf1_pt1_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs0_mf1_pt1_p$aic
)
fs0_mf1_pt1_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs0_mf1_pt1_nb$aic
)
fs0_mf1_pt1_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs0_mf1_pt1_g$aic
)

g_aic <- base::rbind(
  fs0_mf1_pt1_p_aic, 
  fs0_mf1_pt1_nb_aic, 
  fs0_mf1_pt1_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

### 5.1.3 fs 0, mf 2

```{r}
#| label: fs0-mf2-pt0
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 0, membership function = 2, plant type  = 0
# df_dist <- df_dist %>% 
#   dplyr::mutate(
#     class_lag123 = base::ifelse(
#       ac_winterplus_lag123 >= 1134.9,
#       1,
#       0
#     )
#   )

# lag_annual_sum >=480.3 mm
df_dist_fs0_mf2 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 0 & lag_annual_sum >= 480.3
  )

fs0_mf2_pt0_p <- fitdistrplus::fitdist(
  df_dist_fs0_mf2$flowers[df_dist_fs0_mf2$q75_class_cones == 0],
  "pois"
)

fs0_mf2_pt0_nb <- fitdistrplus::fitdist(
  df_dist_fs0_mf2$flowers[df_dist_fs0_mf2$q75_class_cones == 0],
  "nbinom"
)

fs0_mf2_pt0_g <- fitdistrplus::fitdist(
  df_dist_fs0_mf2$flowers[df_dist_fs0_mf2$q75_class_cones == 0],
  "geom"
)

plot(fs0_mf2_pt0_p)
plot(fs0_mf2_pt0_nb)
plot(fs0_mf2_pt0_g)

summary(fs0_mf2_pt0_p)
summary(fs0_mf2_pt0_nb)
summary(fs0_mf2_pt0_g)

# get aic values for each distribution
fs0_mf2_pt0_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs0_mf2_pt0_p$aic
)
fs0_mf2_pt0_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs0_mf2_pt0_nb$aic
)
fs0_mf2_pt0_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs0_mf2_pt0_g$aic
)

g_aic <- base::rbind(
  fs0_mf2_pt0_p_aic, 
  fs0_mf2_pt0_nb_aic, 
  fs0_mf2_pt0_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

```{r}
#| label: fs0-mf2-pt1
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 0, membership function = 2, plant type  = 1

# lag_annual_sum >480.3 mm
df_dist_fs0_mf2 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 0 & lag_annual_sum >= 480
  )


fs0_mf2_pt1_p <- fitdistrplus::fitdist(
  df_dist_fs0_mf2$flowers[df_dist_fs0_mf2$q75_class_cones == 1],
  "pois"
)

fs0_mf2_pt1_nb <- fitdistrplus::fitdist(
  df_dist_fs0_mf2$flowers[df_dist_fs0_mf2$q75_class_cones == 1],
  "nbinom"
)

fs0_mf2_pt1_g <- fitdistrplus::fitdist(
  df_dist_fs0_mf2$flowers[df_dist_fs0_mf2$q75_class_cones == 1],
  "geom"
)

plot(fs0_mf2_pt1_p)
plot(fs0_mf2_pt1_nb)
plot(fs0_mf2_pt1_g)

summary(fs0_mf2_pt1_p)
summary(fs0_mf2_pt1_nb)
summary(fs0_mf2_pt1_g)

# get aic values for each distribution
fs0_mf2_pt1_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs0_mf2_pt1_p$aic
)
fs0_mf2_pt1_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs0_mf2_pt1_nb$aic
)
fs0_mf2_pt1_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs0_mf2_pt1_g$aic
)

g_aic <- base::rbind(
  fs0_mf2_pt1_p_aic, 
  fs0_mf2_pt1_nb_aic, 
  fs0_mf2_pt1_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

### 5.1.4 fs 1, mf 0

```{r}
#| label: fs1-mf0-pt0
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 1, membership function = 0, plant type  = 0
df_dist_fs1_mf0 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 1 & lag_annual_sum <= 412.8
  )

df_dist_fs1_mf0_box <- ggplot(data = df_dist_fs1_mf0) +
  geom_boxplot(aes(x = as.factor(lag_annual_sum), y = flowers))
    
# fuzzy set = 1, membership function = 0, plant type  = 0
test <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 1)

fs1_mf0_pt0_p <- fitdistrplus::fitdist(
  df_dist_fs1_mf0$flowers[df_dist_fs1_mf0$q75_class_cones == 0],
  "pois"
)

fs1_mf0_pt0_nb <- fitdistrplus::fitdist(
  df_dist_fs1_mf0$flowers[df_dist_fs1_mf0$q75_class_cones == 0],
  "nbinom"
)

fs1_mf0_pt0_g <- fitdistrplus::fitdist(
  df_dist_fs1_mf0$flowers[df_dist_fs1_mf0$q75_class_cones == 0],
  "geom"
)

plot(fs1_mf0_pt0_p)
plot(fs1_mf0_pt0_nb)
plot(fs1_mf0_pt0_g)

summary(fs1_mf0_pt0_p)
summary(fs1_mf0_pt0_nb)
summary(fs1_mf0_pt0_g)

# get aic values for each distribution
fs1_mf0_pt0_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs1_mf0_pt0_p$aic
)
fs1_mf0_pt0_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs1_mf0_pt0_nb$aic
)
fs1_mf0_pt0_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs1_mf0_pt0_g$aic
)

g_aic <- base::rbind(
  fs1_mf0_pt0_p_aic, 
  fs1_mf0_pt0_nb_aic, 
  fs1_mf0_pt0_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

```{r}
#| label: fs1-mf0-pt1
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 1, membership function = 0, plant type = 1
df_dist_fs1_mf0 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 1 & lag_annual_sum <= 412.8
  )

fs1_mf0_pt1_p <- fitdistrplus::fitdist(
  df_dist_fs1_mf0$flowers[df_dist_fs1_mf0$q75_class_cones == 1],
  "pois"
)

fs1_mf0_pt1_nb <- fitdistrplus::fitdist(
  df_dist_fs1_mf0$flowers[df_dist_fs1_mf0$q75_class_cones == 1],
  "nbinom"
)

fs1_mf0_pt1_g <- fitdistrplus::fitdist(
  df_dist_fs1_mf0$flowers[df_dist_fs1_mf0$q75_class_cones == 1],
  "geom"
)

plot(fs1_mf0_pt1_p)
plot(fs1_mf0_pt1_nb)
plot(fs1_mf0_pt1_g)

summary(fs1_mf0_pt1_p)
summary(fs1_mf0_pt1_nb)
summary(fs1_mf0_pt1_g)

# get aic values for each distribution
fs1_mf0_pt1_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs1_mf0_pt1_p$aic
)
fs1_mf0_pt1_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs1_mf0_pt1_nb$aic
)
fs1_mf0_pt1_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs1_mf0_pt1_g$aic
)

g_aic <- base::rbind(
  fs1_mf0_pt1_p_aic, 
  fs1_mf0_pt1_nb_aic, 
  fs1_mf0_pt1_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

### 5.1.5 fs 1, mf 1

```{r}
#| label: fs1-mf1-pt0
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 1, membership function = 1, plant type  = 0
df_dist_fs1_mf1 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 1 & lag_annual_sum == 458.3
  )

fs1_mf1_pt0_p <- fitdistrplus::fitdist(
  df_dist_fs1_mf1$flowers[df_dist_fs1_mf1$q75_class_cones == 0],
  "pois"
)

fs1_mf1_pt0_nb <- fitdistrplus::fitdist(
  df_dist_fs1_mf1$flowers[df_dist_fs1_mf1$q75_class_cones == 0],
  "nbinom"
)

fs1_mf1_pt0_g <- fitdistrplus::fitdist(
  df_dist_fs1_mf1$flowers[df_dist_fs1_mf1$q75_class_cones == 0],
  "geom"
)

plot(fs1_mf1_pt0_p)
plot(fs1_mf1_pt0_nb)
plot(fs1_mf1_pt0_g)

summary(fs1_mf1_pt0_p)
summary(fs1_mf1_pt0_nb)
summary(fs1_mf1_pt0_g)

# get aic values for each distribution
fs1_mf1_pt0_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs1_mf1_pt0_p$aic
)
fs1_mf1_pt0_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs1_mf1_pt0_nb$aic
)
fs1_mf1_pt0_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs1_mf1_pt0_g$aic
)

g_aic <- base::rbind(
  fs1_mf1_pt0_p_aic, 
  fs1_mf1_pt0_nb_aic, 
  fs1_mf1_pt0_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

```{r}
#| label: fs1-mf1-pt1
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 1, membership function = 1, plant type  = 1
df_dist_fs1_mf1 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 1 & lag_annual_sum == 458.3
  )

fs1_mf1_pt1_p <- fitdistrplus::fitdist(
  df_dist_fs1_mf1$flowers[df_dist_fs1_mf1$q75_class_cones == 1],
  "pois"
)

fs1_mf1_pt1_nb <- fitdistrplus::fitdist(
  df_dist_fs1_mf1$flowers[df_dist_fs1_mf1$q75_class_cones == 1],
  "nbinom"
)

fs1_mf1_pt1_g <- fitdistrplus::fitdist(
  df_dist_fs1_mf1$flowers[df_dist_fs1_mf1$q75_class_cones == 1],
  "geom"
)

plot(fs1_mf1_pt1_p)
plot(fs1_mf1_pt1_nb)
plot(fs1_mf1_pt1_g)

summary(fs1_mf1_pt1_p)
summary(fs1_mf1_pt1_nb)
summary(fs1_mf1_pt1_g)

# get aic values for each distribution
fs1_mf1_pt1_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs1_mf1_pt1_p$aic
)
fs1_mf1_pt1_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs1_mf1_pt1_nb$aic
)
fs1_mf1_pt1_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs1_mf1_pt1_g$aic
)

g_aic <- base::rbind(
  fs1_mf1_pt1_p_aic, 
  fs1_mf1_pt1_nb_aic, 
  fs1_mf1_pt1_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

### 5.1.6 fs 1, mf 2

```{r}
#| label: fs1-mf2-pt0
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 1, membership function = 2, plant type  = 0
df_dist_fs1_mf2 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 1 & lag_annual_sum == 492.7
)

fs1_mf2_pt0_p <- fitdistrplus::fitdist(
  df_dist_fs1_mf2$flowers[df_dist_fs1_mf2$q75_class_cones == 0],
  "pois"
)

fs1_mf2_pt0_nb <- fitdistrplus::fitdist(
  df_dist_fs1_mf2$flowers[df_dist_fs1_mf2$q75_class_cones == 0],
  "nbinom"
)

fs1_mf2_pt0_g <- fitdistrplus::fitdist(
  df_dist_fs1_mf2$flowers[df_dist_fs1_mf2$q75_class_cones == 0],
  "geom"
)

plot(fs1_mf2_pt0_p)
plot(fs1_mf2_pt0_nb)
plot(fs1_mf2_pt0_g)

summary(fs1_mf2_pt0_p)
summary(fs1_mf2_pt0_nb)
summary(fs1_mf2_pt0_g)

# get aic values for each distribution
fs1_mf2_pt0_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs1_mf2_pt0_p$aic
)
fs1_mf2_pt0_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs1_mf2_pt0_nb$aic
)
fs1_mf2_pt0_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs1_mf2_pt0_g$aic
)

g_aic <- base::rbind(
  fs1_mf2_pt0_p_aic, 
  fs1_mf2_pt0_nb_aic, 
  fs1_mf2_pt0_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

```{r}
#| label: fs1-mf2-pt1
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 1, membership function = 2, plant type  = 1
df_dist_fs1_mf2 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 1 & lag_annual_sum == 492.7
)

fs1_mf2_pt1_p <- fitdistrplus::fitdist(
  df_dist_fs1_mf2$flowers[df_dist_fs1_mf2$q75_class_cones == 1],
  "pois"
)

fs1_mf2_pt1_nb <- fitdistrplus::fitdist(
  df_dist_fs1_mf2$flowers[df_dist_fs1_mf2$q75_class_cones == 1],
  "nbinom"
)

fs1_mf2_pt1_g <- fitdistrplus::fitdist(
  df_dist_fs1_mf2$flowers[df_dist_fs1_mf2$q75_class_cones == 1],
  "geom"
)

plot(fs1_mf2_pt1_p)
plot(fs1_mf2_pt1_nb)
plot(fs1_mf2_pt1_g)

summary(fs1_mf2_pt1_p)
summary(fs1_mf2_pt1_nb)
summary(fs1_mf2_pt1_g)

# get aic values for each distribution
fs1_mf2_pt1_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs1_mf2_pt1_p$aic
)
fs1_mf2_pt1_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs1_mf2_pt1_nb$aic
)
fs1_mf2_pt1_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs1_mf2_pt1_g$aic
)

g_aic <- base::rbind(
  fs1_mf2_pt1_p_aic, 
  fs1_mf2_pt1_nb_aic, 
  fs1_mf2_pt1_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

### 5.1.7 fs 1, mf 3

```{r}
#| label: fs1-mf3-pt0
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 1, membership function = 3, plant type  = 0
df_dist_fs1_mf3 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 1 & lag_annual_sum >= 518.3
)

fs1_mf3_pt0_p <- fitdistrplus::fitdist(
  df_dist_fs1_mf3$flowers[df_dist_fs1_mf3$q75_class_cones == 0],
  "pois"
)

fs1_mf3_pt0_nb <- fitdistrplus::fitdist(
  df_dist_fs1_mf3$flowers[df_dist_fs1_mf3$q75_class_cones == 0],
  "nbinom"
)

fs1_mf3_pt0_g <- fitdistrplus::fitdist(
  df_dist_fs1_mf3$flowers[df_dist_fs1_mf3$q75_class_cones == 0],
  "geom"
)

plot(fs1_mf3_pt0_p)
plot(fs1_mf3_pt0_nb)
plot(fs1_mf3_pt0_g)

summary(fs1_mf3_pt0_p)
summary(fs1_mf3_pt0_nb)
summary(fs1_mf3_pt0_g)

# get aic values for each distribution
fs1_mf3_pt0_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs1_mf3_pt0_p$aic
)
fs1_mf3_pt0_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs1_mf3_pt0_nb$aic
)
fs1_mf3_pt0_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs1_mf3_pt0_g$aic
)

g_aic <- base::rbind(
  fs1_mf3_pt0_p_aic, 
  fs1_mf3_pt0_nb_aic, 
  fs1_mf3_pt0_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

```{r}
#| label: fs1-mf3-pt1
#| include: true
#| echo: false
#| collapse: false
#| warning: false

# load libraries
library(dplyr)

# fuzzy set = 1, membership function = 3, plant type  = 1
df_dist_fs1_mf3 <- flower_current_q_all_2017 %>% 
  dplyr::filter(
    fuzzy == 1 & lag_annual_sum >= 518.3
)

fs1_mf3_pt1_p <- fitdistrplus::fitdist(
  df_dist_fs1_mf3$flowers[df_dist_fs1_mf3$q75_class_cones == 1],
  "pois"
)

fs1_mf3_pt1_nb <- fitdistrplus::fitdist(
  df_dist_fs1_mf3$flowers[df_dist_fs1_mf3$q75_class_cones == 1],
  "nbinom"
)

fs1_mf3_pt1_g <- fitdistrplus::fitdist(
  df_dist_fs1_mf3$flowers[df_dist_fs1_mf3$q75_class_cones == 1],
  "geom"
)

plot(fs1_mf3_pt1_p)
plot(fs1_mf3_pt1_nb)
plot(fs1_mf3_pt1_g)

summary(fs1_mf3_pt1_p)
summary(fs1_mf3_pt1_nb)
summary(fs1_mf3_pt1_g)

# get aic values for each distribution
fs1_mf3_pt1_p_aic <- base::data.frame(
  distribution = "poisson", 
  AIC = fs1_mf3_pt1_p$aic
)
fs1_mf3_pt1_nb_aic <- base::data.frame(
  distribution = "negative binomial", 
  AIC = fs1_mf3_pt1_nb$aic
)
fs1_mf3_pt1_g_aic <- base::data.frame(
  distribution = "geometric",
  AIC = fs1_mf3_pt1_g$aic
)

g_aic <- base::rbind(
  fs1_mf3_pt1_p_aic, 
  fs1_mf3_pt1_nb_aic, 
  fs1_mf3_pt1_g_aic
)

knitr::kable(
  g_aic, 
  caption = "AIC values of the three distribution tested"
)

```

## 5.2 Plot all flower count distributions

```{r}
# load libraries
library(ggplot2)
library(ggpubr)
library(tidyr)


# fs0_mf1
g_n = 0:30

g_df_fs0_mf1 <- data.frame(
  events = g_n, 
  pt0 = stats::dpois(x = g_n, lambda = 0.071),
  pt1 = stats::dpois(x = g_n, lambda = 0.111)
)

g_df_fs0_mf1 <- g_df_fs0_mf1 %>% 
  tidyr::pivot_longer(
    -c(events),
    names_to = "plant_type",
    values_to = "density"
  )

p_fs0_mf1 <- ggplot(
  data = g_df_fs0_mf1, 
  mapping = aes(
    x = factor(events),
    y = density,
    fill = plant_type)) +
  geom_col(position = "dodge") +
  labs(x = "Flowers (#)", y = "Density") +
  theme_classic() +
  scale_fill_manual(
    name = "Flower producer type:",
    # values = c("#56B4E9", "#E69F00"),
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("low", "high")
  ) +
  scale_x_discrete(breaks = seq(0, 30, 5)) +
  ggtitle("(a)") +
  theme(
    text = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  )


# fs0_mf2
g_df_fs0_mf2 <- data.frame(
  events = g_n, 
  pt0 = stats::dgeom(x = g_n, prob = 0.673),
  pt1 = stats::dgeom(x = g_n, prob = 0.288)
)

g_df_fs0_mf2 <- g_df_fs0_mf2 %>% 
  tidyr::pivot_longer(
    -c(events),
    names_to = "plant_type",
    values_to = "density"
  )

p_fs0_mf2 <- ggplot(
  data = g_df_fs0_mf2, 
  mapping = aes(
    x = factor(events),
    y = density,
    fill = plant_type)) +
  geom_col(position = "dodge") +
  labs(x = "Flowers (#)", y = "Density") +
  theme_classic() +
  scale_fill_manual(
    name = "Flower producer type:",
    # values = c("#56B4E9", "#E69F00"),
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("low", "high")
  ) +
  scale_x_discrete(breaks = seq(0, 30, 5)) +
  ggtitle("(b)") +
  theme(
    text = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  )

# fs1_mf0
g_df_fs1_mf0 <- data.frame(
  events = g_n, 
  pt0 = stats::dgeom(x = g_n, prob = 0.618),
  pt1 = stats::dgeom(x = g_n, prob = 0.415)
)

g_df_fs1_mf0 <- g_df_fs1_mf0 %>% 
  tidyr::pivot_longer(
    -c(events),
    names_to = "plant_type",
    values_to = "density"
  )

p_fs1_mf0 <- ggplot(
  data = g_df_fs1_mf0, 
  mapping = aes(
    x = factor(events),
    y = density,
    fill = plant_type)) +
  geom_col(position = "dodge") +
  labs(x = "Flowers (#)", y = "Density") +
  theme_classic() +
  scale_fill_manual(
    name = "Flower producer:",
    # values = c("#56B4E9", "#E69F00"),
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("low", "high")
  ) +
  scale_x_discrete(breaks = seq(0, 30, 5)) +
  ggtitle("(c)") + 
    theme(
    text = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  )


# fs1_mf1
g_df_fs1_mf1 <- data.frame(
  events = g_n, 
  pt0 = stats::dpois(x = g_n, lambda = 1.214),
  pt1 = stats::dgeom(x = g_n, prob = 0.25)
)

g_df_fs1_mf1 <- g_df_fs1_mf1 %>% 
  tidyr::pivot_longer(
    -c(events),
    names_to = "plant_type",
    values_to = "density"
  )

p_fs1_mf1 <- ggplot(
  data = g_df_fs1_mf1, 
  mapping = aes(
    x = factor(events),
    y = density,
    fill = plant_type)) +
  geom_col(position = "dodge") +
  labs(x = "Flowers (#)", y = "Density") +
  theme_classic() +
  scale_fill_manual(
    name = "Flower producer:",
    # values = c("#56B4E9", "#E69F00"),
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("low", "high")
  ) +
  scale_x_discrete(breaks = seq(0, 30, 5)) +
  ggtitle("(d)") +
    theme(
    text = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  )

# fs1_mf2
g_df_fs1_mf2 <- data.frame(
  events = g_n, 
  pt0 = stats::dgeom(x = g_n, prob = 0.203),
  pt1 = stats::dnbinom(x = g_n, size = 3.254, mu = 8.443)
)

g_df_fs1_mf2 <- g_df_fs1_mf2 %>% 
  tidyr::pivot_longer(
    -c(events),
    names_to = "plant_type",
    values_to = "density"
  )

p_fs1_mf2 <- ggplot(
  data = g_df_fs1_mf2, 
  mapping = aes(
    x = factor(events),
    y = density,
    fill = plant_type)) +
  geom_col(position = "dodge") +
  labs(x = "Flowers (#)", y = "Density") +
  theme_classic() +
  scale_fill_manual(
    name = "Flower producer:",
    # values = c("#56B4E9", "#E69F00"),
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("low", "high")
  ) +
  scale_x_discrete(breaks = seq(0, 30, 5)) +
  ggtitle("(e)") +
    theme(
    text = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  )


# fs1_mf3
g_df_fs1_mf3 <- data.frame(
  events = g_n, 
  pt0 = stats::dgeom(x = g_n, prob = 0.205),
  pt1 = stats::dnbinom(x = g_n, size = 6.274, mu = 10.100)
)

g_df_fs1_mf3 <- g_df_fs1_mf3 %>% 
  tidyr::pivot_longer(
    -c(events),
    names_to = "plant_type",
    values_to = "density"
  )

p_fs1_mf3 <- ggplot(
  data = g_df_fs1_mf3, 
  mapping = aes(
    x = factor(events),
    y = density,
    fill = plant_type)) +
  geom_col(position = "dodge") +
  labs(x = "Flowers (#)", y = "Density") +
  theme_classic() +
  scale_fill_manual(
    name = "Flower producer:",
    # values = c("#56B4E9", "#E69F00"),
    values = c("#0073C2FF", "#EFC000FF"),
    labels = c("low", "high")
  ) +
  scale_x_discrete(breaks = seq(0, 30, 5)) +
  ggtitle("(f)") +
    theme(
    text = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position = "none"
  )


# legend
# p_legend <- ggplot() +
#   geom_col(
#     data = g_df_fs1_mf3, 
#     mapping = aes(
#       x = as.numeric(events),
#       y = as.numeric(density),
#       fill = plant_type),
#     position = "dodge") +
#   # labs(x = "# of flowers", y = "density") +
#   # theme_classic() +
#   scale_fill_manual(
#     name = "Flower producer class",
#     # values = c("#56B4E9", "#E69F00"),
#     values = c("#0073C2FF", "#EFC000FF"),
#     labels = c("poor", "good")
#   ) +
#   # scale_x_discrete(breaks = seq(0, 30, 5)) +
#   ggplot2::xlim(0, 0) +
#   ggplot2::ylim(0, 0) +
#   ggtitle(" ") +
#   # theme(
#   #   text = element_text(size = 12, face = "bold"),
#   #   axis.text    = element_text(size = 12),
#   #   axis.title   = element_text(size = 12, face = "bold"),
#   #   legend.text  = element_text(size = 12, face = "plain"),
#   #   legend.title = element_text(size = 12, face = "bold")
#   #   # legend.position = "none"
#   # ) +
#   # ggplot2::lims(x = c(0, 0), y = c(0, 0)) +
#   # ggplot2::coord_cartesian(xlim = c(-1, -1), ylim = c(-1, -1)) +
#   theme_void() +
#   theme(
#     legend.position = c(0.5, 0.5),
#     # legend.key.size = unit(1, "cm"),
#     legend.text  = element_text(size = 12, face = "plain"),
#     legend.title = element_text(size = 12, face = "bold")
#   ) +
#   guides(colour = guide_legend(override.aes = list(size = 8)))

# combine all subplots
g_parr <- ggpubr::ggarrange(
  p_fs0_mf1,
  p_fs0_mf2,
  p_fs1_mf0,
  p_fs1_mf1,
  p_fs1_mf2,
  p_fs1_mf3,
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom"
)

g_parr

# save plot
ggplot2::ggsave(
  plot = g_parr,
  filename = base::paste0("flower_distr_dpi300_6plots.tiff"),
  path = "./data/plots",
  width = 3.31 * 2,
  height = 3.31 * 1.5,
  bg = "white",
  dpi = 300
)


```

```{r}
#| label: fuzzy-paper-plot
#| include: true
#| echo: false
#| collapse: false
#| warning: false


# load libraries
# library(ggplot2)
# library(ggpubr)
# 
# p3f_copy <- p3f + ggtitle("(a)")
# p_fs1_mf0_copy <- p_fs1_mf0 + 
#   ggtitle("(b)") + 
#   ggplot2::theme(
#     text         = element_text(size = 14, face = "bold"),
#     axis.text    = element_text(size = 14),
#     axis.title   = element_text(size = 14, face = "bold"),
#     legend.text  = element_text(size = 14),
#     legend.title = element_text(size = 14, face = "bold"),
#   )
# 
# p_fs1_mf3_copy <- p_fs1_mf3 + 
#   ggtitle("(c)") +
#   ggplot2::theme(
#     text         = element_text(size = 14, face = "bold"),
#     axis.text    = element_text(size = 14),
#     axis.title   = element_text(size = 14, face = "bold"),
#     legend.text  = element_text(size = 14),
#     legend.title = element_text(size = 14, face = "bold"),
#   ) 
# 
# g_parr1 <- ggpubr::ggarrange(
#   p3f_copy,
#   ncol = 1,
#   common.legend = TRUE, 
#   legend = "bottom"
# )
# 
# g_parr2 <- ggpubr::ggarrange(
#   p_fs1_mf0_copy,
#   p_fs1_mf3_copy,
#   ncol = 2,
#   common.legend = TRUE, 
#   legend = "bottom"
# )
# 
# g_parr3 <- ggpubr::ggarrange(
#   g_parr1,
#   g_parr2,
#   ncol = 1,
#   nrow = 2
# )
# 
# 
# ggplot2::ggsave(
#   plot = g_parr3,
#   filename = base::paste0("fuzzy_paper_plot.tiff"),
#   path = "./data/plots",
#   width = 3.31 * 2,
#   height = 3.31 * 2,
#   bg = "white",
#   dpi = 1200
# )

```

# 6. Sensitivity analysis

```{r}
#| label: sensitivity-calc
#| include: true
#| echo: true
#| collapse: false
#| warning: false

# load utils
source("utils/pooled_sd.R")
source("utils/read_multiple_files.R")

library(signs)
library(tibble)

# input files
g_path <- "data/sim_outputs/sensitivity/"
g_path_ref <- "data/sim_outputs/sensitivity/sens_current_ref"
g_path_sens <- "data/sim_outputs/sensitivity/sens_current"

# simulation output type
g_sim_type <- "*_metapop.csv"

# name output figure
g_figure_name <- "sensitivity"

# read simulation experiment output files and create df
g_df_ref <- fun_read_files(base::paste0(g_path_ref, "/raw"), g_sim_type)
g_df_sens <- fun_read_files(base::paste0(g_path_sens, "/raw"), g_sim_type)

# read simulation parameters
g_params_ref <- base::list.files(
  path = base::paste0(g_path),
  pattern = "sens_current_ref_values.csv",
  full.names = TRUE
)

g_params_ref <- plyr::ldply(
  g_params_ref,
  utils::read.table,
  sep = " ",
  fill = TRUE,
  header = TRUE
)


# take the parameter values for plotting
g_params_ref <- g_params_ref[, c("sim_id", "param1", "val1")]


g_params_ref_unique <- g_params_ref[!base::duplicated(g_params_ref[ , base::c("param1")]), ]

g_params_ref_unique <- tibble::rowid_to_column(g_params_ref_unique, "id")

# g_params_ref_unique <- g_params_ref_unique[, base::c(1, 3)]

# read simulation parameters
g_params <- base::list.files(
  path = base::paste0(g_path_sens, "/in/sim"),
  pattern = "*_simfile.csv",
  full.names = TRUE
)

g_params <- plyr::ldply(
  g_params,
  utils::read.table,
  sep = " ",
  fill = TRUE,
  header = TRUE
)

# take the parameter values for plotting
g_params <- g_params[, c("sim_id", "param1", "val1")]


g_params_unique <- g_params[!base::duplicated(g_params[ , base::c("param1")]), ]

# g_params$sim_id <- base::seq.int(base::nrow(g_params))
g_params_unique <- tibble::rowid_to_column(g_params_unique, "id")

g_params_unique <- g_params_unique[, base::c(1, 3)]

# calculate metapopulation persistence
g_df_ref_summ <- g_df_ref %>%
  dplyr::group_by(sim_id) %>%
  dplyr::summarise(
    grand_mean_year_max = base::mean(year_max, na.rm = TRUE),
    grand_mean_occu_mean = base::mean(occu_mean, na.rm = TRUE)
  )

g_df_sens_summ <- g_df_sens %>%
  dplyr::group_by(sim_id) %>%
  dplyr::summarise(
    grand_mean_year_max = base::mean(year_max, na.rm = TRUE),
    grand_mean_occu_mean = base::mean(occu_mean, na.rm = TRUE)
  )

g_df_summ <- dplyr::mutate(
  g_df_sens_summ,
  dev_persistence = (grand_mean_year_max - g_df_ref_summ$grand_mean_year_max[1]) / 
      g_df_ref_summ$grand_mean_year_max[1] * 100,
  dev_occur = (grand_mean_occu_mean - g_df_ref_summ$grand_mean_occu_mean[1]) / 
      g_df_ref_summ$grand_mean_occu_mean[1] * 100
)

df <- base::as.data.frame(
  base::matrix(
    base::round(g_df_summ$dev_persistence, 2),
    ncol = 4,
    byrow = TRUE
  )
)

# change column names
base::names(df) <- base::c("\u201310", "\u20135", "+5", "+10")

# change row names
base::row.names(df) <- base::paste(g_params_unique$param1)
```

```{r}
#| label: sensitivity-table-values
#| include: true
#| echo: false
#| collapse: false
#| warning: false 

library(dplyr)
library(knitr)
library(kableExtra)
library(signs)

g_df_params_with_ref <- base::as.data.frame(
  base::matrix(
    signs::signs(g_params_ref_unique$val1, accuracy = 0.001),
    ncol = 1, 
    byrow = TRUE
  )
)

g_df_params <- base::as.data.frame(
  base::matrix(
    signs::signs(g_params$val1, accuracy = 0.001),
    ncol = 4, 
    byrow = TRUE
  )
)

g_df_params_with_ref <- cbind(g_df_params_with_ref, g_df_params) 


# change column names
base::names(g_df_params_with_ref) <- base::c(" ", "\u201310", "\u20135", "+5", "+10")

# change row names
base::row.names(g_df_params_with_ref) <- base::paste(g_params_unique$param1)


g_tb_params <- g_df_params_with_ref %>%  
  kableExtra::kbl(caption = "Table xx Parameter variation values" , booktabs = TRUE, align = "r") %>% 
  kableExtra::add_header_above(base:::c("Parameter code", "Reference", "Variation (%)" = 4), align = "c", line = F) %>%
  kable_styling(full_width = TRUE, latex_options = c("striped", "hold_position")) %>% 
  kableExtra::row_spec(7, color = c(rep("red", 4))) %>%
  kableExtra::row_spec(8, color = c(rep("red", 4))) %>%
  kableExtra::row_spec(9, color = c(rep("red", 4))) %>%
  kableExtra::row_spec(18, color = c(rep("red", 4))) %>%
  kableExtra::row_spec(19, color = c(rep("red", 4))) %>%
  kableExtra::row_spec(20, color = c("red")) %>%
  kableExtra::row_spec(21, color = c("red")) %>%
  kableExtra::row_spec(22, color = c("red")) %>%
  kableExtra::row_spec(24, color = c(rep("red", 4))) %>%
  kableExtra::row_spec(33, color = c(rep("red", 4))) %>%
  kableExtra::row_spec(34, color = c(rep("red", 4))) %>%
  kableExtra::column_spec(2, color = "black") %>% 
  kableExtra::column_spec(3, color = c(rep("black", 6), rep("red", 3), rep("black", 8), rep("red", 2), "black", "red", "black", "black", "red", rep("black", 8), rep("red", 2), rep("black", 20))) %>% 
  kableExtra::column_spec(4, color = c(rep("black", 6), rep("red", 3), rep("black", 8), rep("red", 2), rep("black", 4), "red", rep("black", 8), rep("red", 2), rep("black", 20))) %>% 
    kableExtra::column_spec(5, color = c(rep("black", 6), rep("red", 3), rep("black", 8), rep("red", 2), "black", "red", "black", "black", "red", rep("black", 8), rep("red", 2), rep("black", 20))) %>%  
  kableExtra::column_spec(6, color = c(rep("black", 6), rep("red", 3), rep("black", 8), rep("red", 2), "red", "red", "red", "black", "red", rep("black", 8), rep("red", 2), rep("black", 20)))

g_tb_params

# write csv file
# utils::write.csv(x = g_df_params_with_ref, file = "sensitivity_values.csv")
```

```{r}
#| label: sensitivity-table-results
#| include: true
#| echo: false
#| collapse: false
#| warning: false 

library(dplyr)
library(knitr)
library(kableExtra)
library(webshot2)

g_tb <- df %>%  
  kableExtra::kbl(
    caption = "Table xx Local sensitivity analysis" , 
    booktabs = TRUE, 
    align = "r"
  ) %>% 
  kableExtra::add_header_above(
    base:::c("Parameter code", "Variation (%)" = 4),
    align = "c",
    line = FALSE
  ) %>% 
  kableExtra::column_spec(
    column = 2, 
    background = base::ifelse(
      df$`–10` < -20,
      "#EFC000FF",
      base::ifelse(
        df$`–10` > 20,
        "#0073C2FF",
        "white")
    )
  ) %>% 
  kableExtra::column_spec(
    column = 3, 
    background = base::ifelse(
      df$`–5` < -10,
      "#EFC000FF", 
      base::ifelse(
        df$`–5` > 10, 
        "#0073C2FF", 
        "white")
    )
  ) %>%
  kableExtra::column_spec(
    column = 4, 
    background = base::ifelse(
      df$`+5` < -10,
      "#EFC000FF", 
      base::ifelse(
        df$`+5` > 10,
        "#0073C2FF",
        "white")
    )
  ) %>% 
  kableExtra::column_spec(
    column = 5,
    background = base::ifelse(
      df$`+10` < -20,
      "#EFC000FF", 
      base::ifelse(
        df$`+10` > 20, 
        "#0073C2FF", 
        "white")
    )
  ) %>% 
  kable_styling(latex_options = c("striped", "hold_position"))

g_tb

# kableExtra::save_kable(g_tb, file = "./mtcars.html")
# webshot2::webshot("./mtcars.html", "./mtcars.pdf")
# 
# # write csv file
# utils::write.csv(x = df, file = "sensitivity_results.csv")

# kableExtra::save_kable(
#   x = g_tb,
#   # file = "./data/sim_outputs/sensitivity/tables/sensitivity_results.pdf"
#     file = "./test.pdf"
# )

```

```{r}

## exort table as latex code. There are two different packages that do this work
# 
# library(Hmisc)
# 
# Hmisc::latex(df, file="")            # If you want all the data
# latex(describe(d), file="")  # If you just want a summary

# library(xtable)
# 
# xtable::xtable(df)

```
